### 2020-10-14 paper-ffs.tex
init from MT-paper-FSs

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage[numbers]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[warn]{textcomp}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{commath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[super]{nth}
\usepackage{url}

\def\UrlBreaks{\do\/\do-}

% tikz
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
% tikz styles
\tikzstyle{user} = [circle, thick, draw, minimum size=2.5cm]
\tikzstyle{item} = [rectangle, thick, draw, minimum width=3cm, minimum height=1.5cm]
\tikzstyle{arrow} = [->, draw, thick]
\tikzstyle{bi-arrow} = [<->, draw, thick]
\tikzstyle{line} = [-, draw, thick]

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\emergencystretch=8em

\begin{document}

\title{Community Embeddings for Facebook Friend Suggestions\\
    \thanks{The reported study was partially supported by RFBR grant \textnumero 20-07-00958. The paper was prepared within the framework of the HSE University Project Group Competition 2020-2022.}
}

\author{
    \IEEEauthorblockN{Anton I. N. Begehr}
    \IEEEauthorblockA{
        \textit{Graduate School of Business}\\
        \textit{National Research University Higher School of Economics}\\
        Moscow, Russia\\
        a.begehr@fu-berlin.de\\
    }
    \and
    \IEEEauthorblockN{Prof. Dr. Petr Panfilov}
    \IEEEauthorblockA{
        \textit{Graduate School of Business}\\
        \textit{National Research University Higher School of Economics}\\
        Moscow, Russia\\
        ppanfilov@hse.ru\\
    }
}

\maketitle

\begin{abstract}
    Graphs, such as social networks, emerge naturally from various real-world situations. Recently, graph embedding methods have gained traction in data science research.
    Recommender systems are used in a wide range of business applications and are essential for online e-business models to survive and strive in the contemporary market. Using graph embeddings for recommendation tasks can improve upon recommender systems because of data compression, the feature vector format, and sub-quadratic time complexity on computing embeddings.
    Graph and community embeddings generated with ComE BGMM+VI are used to build a recommender system for friend suggestions. ComE BGMM+VI is an alteration of the community embeddings algorithm ComE. ComE BGMM+VI applies a Bayesian Gaussian mixture model and variational inference for community embedding and detection.
    Recommendations are evaluated by the top-N hit-rate for users with at least 50 friends. A friend suggestions recommender system with a top-10 leave-one-out hit-rate of 43.6\% and run-time optimized hit-rate of 32.9\% is presented.
\end{abstract}

\begin{IEEEkeywords}
    graph, embedding, community embedding, ComE, recommender systems, friend suggestions
\end{IEEEkeywords}

\section{Introduction}

Graphs, such as social networks, knowledge graphs, content-rating graphs, and communication networks emerge naturally from various real-world situations. Analyzing these graphs leads to findings and understanding of the underlying structures, coherences, and dependencies. Recently, methods for embedding graph's nodes into lower-dimensional, Euclidean spaces, called graph embeddings, have gained traction in multiple areas of data science research \cite{Goyal_2018}.

Due to the rapid growth of the internet and data accumulation, recommender systems are essential for e-business and online business models to survive and strive in the contemporary market \cite{Polatidis2013RecommenderST}. Modern recommender systems need to take into account the huge amounts of data generated at all times in big data systems around the world and improve recommendations instead of failing under the thrust of big data overload.

Utilizing graph embeddings for recommendation tasks, has recently gained research interest \cite{Palumbo2018AnEC, Palumbo2018KnowledgeGE, GradGyenge2017GraphEB, Sathish2019GraphEB}. The advantages of graph embeddings include data compression and the feature vector format \cite{Godec2018}. Given these advantages and provided competitive results, graph embeddings have the possibility of greatly improving upon graph-based use-cases like recommender systems.

Community Embeddings, in addition to embedding a graph's nodes through first- and second-order proximity, also preserve higher-order proximity by embedding clusters present in the graph data. The graph and community embedding algorithm ComE aims to preserve first-, second- and higher-order proximity by embedding a graph's nodes and communities \cite{ComE}.

This work specifically examines community embeddings for friend suggestion recommender systems and evaluates recommendations on social network graph data for the use-case of friend suggestions. Graph and community embeddings generated with ComE BGMM+VI are used to develop a friend suggestions recommender system based on the shortest distances between nodes in the embedding. Recommendations are evaluated by the top-$N$ recommendations hit-rate of test edges. A friend suggestions recommender system with a top-10 leave-one-out hit-rate of 43.6\% and run-time optimized 32.9\% is presented.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.48\textwidth]{images/fb/graph_d4.png}
    \caption{\textit{Social Circles: Facebook} graph plotted with networkx's spring layout and node colors representing community membership obtained by ComE BGMM+VI with $D=4$ \cite{networkx}.}
    \label{fig:fb_graph}
\end{figure}

\section{Friend Suggestions}

Recommender Systems are eagerly researched in academia and widely deployed in real-world business applications. Most contemporary technology companies heavily rely on recommender systems to drive usage of their services and consumption of their content. Users, in turn, rely on recommender systems to find what they want, need, save searching time, and what they might not have found otherwise. State-of-the-art recommender systems provide a competitive advantage desperately needed by online services. Companies heavily relying on recommender systems include YouTube, Amazon, Netflix, and many more \cite{Rocca19}.

Recommender systems are built on top of user-item interaction. Friend suggestions are a simpler type of recommender system. For friend suggestions no distinction is made between users and items. The entity user is both the subject and object of recommendation. This results in a simple data model, which can be used for structural friend suggestions, as shown in Fig. \ref{fig:data_model}.

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}[node distance=6cm]
        \node (user) [user] {User};
        \path [arrow] (user) edge[loop right] node[right]{relate} (user);
    \end{tikzpicture}
    \caption{Minimal data model supporting friend suggestions.}
    \label{fig:data_model}
\end{figure}

The two main methods used for recommender systems are collaborative filtering and content-based filtering. Collaborative filtering is based on the assumption that users like items similar to other items they like and items that are liked by other users with similar tastes \cite{Grover17, Su2009ASO}. Content-based filtering considers user and item attributes, instead of solely interactions \cite{Rocca19, Lops2011SurveyCB}.

In this paper, a method of generating friend suggestions based solely on graph data structure, and not on the node or edge attributes, is presented. The graph of users and friendships is embedded to a lower-dimensional space with the community embedding algorithm ComE BGMM+VI \cite{ComE, ComE+, ComE_BGMM_GH}. The approach can be considered a variant of collaborative filtering with the Euclidean distance of user embeddings and user community membership as measures for user similarity. The proposed advantage of using such an approach is that community embeddings optimize first-, second-, and higher-order proximity between users in the node embedding.

\subsection{Evaluation}

To determine the effectivity of friend suggestions, the generated friend suggestions must be properly validated. In this paper, the top-$N$ approach to evaluating recommender systems with hit-rate as the evaluation metric is chosen.

A user's top-$N$ recommendations is a list of $N$ items to be recommended to a specific user. To evaluate recommender systems with the top-$N$ and hit-rate approach, initially, the dataset is split into train and test data. For each testing user, one relation is left out, according to the leave-one-out method and the model is trained on the remaining training dataset. Once the model is trained, a list of top-$N$ recommendations is generated for each testing user. If the item corresponding to the user is in the user's top-$N$ list, a hit is counted, otherwise, a miss is counted. The hit-rate is defined as the total number of hits divided by the number of testing users.

Utilizing the top-$N$ approach with hit-rate to evaluate recommender systems is advantageous to evaluating recommender systems by a link prediction approach since the top-$N$ approach is more realistic in comparison to actual recommender system use-cases. When you open Netflix, Amazon, YouTube, or Facebook one or multiple top-$N$ recommendation lists are generated and displayed. If you click on an item and buy, watch, or befriend, that is considered a hit, otherwise, a miss.

The hit-rate metric on top-$N$ recommendations provides a realistic option of evaluating recommender systems and is often used with the leave-one-out (LOO) strategy in literature \cite{Cremonesi2010PerformanceOR, Palumbo2017entity2rec, Zhao2019ATS, He2017NeuralCF}.

\section{Algorithm}

The proposed algorithm for generating friend suggestions using community embeddings is detailed and evaluated on time complexity. The initially quadratic runtime of generating friend suggestions for all users is then reduced by a factor of $K$ by utilizing a node's community membership. The two resulting algorithms are presented.

Algorithm \ref{alg:friend_sug} describes in pseudocode how recommendations are computed from node embeddings generated by ComE BGMM+VI. The terms friend suggestions and social recommendations are used interchangeably.

\begin{algorithm}
    \caption{Top-$N$ Social Recommendations based on Node Embeddings}
    \label{alg:friend_sug}
    \begin{algorithmic}[1]
        \Require{graph $G = (V, E)$, maximum number of communities K, number of walks $\gamma$, walk length $\ell$, window size $\zeta$, representation size $D$, negative context size $m$, parameters ($\alpha$, $\beta$), number of recommendations $N$.}
        \Ensure{Top-$N$ recommendations for all nodes $R$.}
        % run ComE BGMM+VI
        \State $\Phi, \Phi', \Pi, (\Psi,\Sigma) \leftarrow ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$
        % find friends
        \For{$(v, v') \in E$}
        \State $F_v \leftarrow F_v \cup \{v'\}$
        \State $F_{v'} \leftarrow F_{v'} \cup \{v\}$
        \EndFor
        % for each node
        \For{$v \in V$}
        % generate ascending recommendation list for each node
        \State $R_v \leftarrow SortedDict(size=N)$
        \For{$v' \in V \land v' \notin F_v \land v' \neq v$}
        \State $d \leftarrow \|\phi_v - \phi_{v'}\|$
        \State $R_v[d] \leftarrow v'$
        \EndFor
        % get Top-N recommendations for each node
        \State $R_v  \leftarrow R_v.values()$
        \EndFor
    \end{algorithmic}
\end{algorithm}

The function $ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ is the ComE BGMM+VI community embedding algorithm. It returns the node embedding $\Phi$, context embedding $\Phi'$, community assignment $\Pi$, and community embedding $(\Psi,\Sigma)$. ComE BGMM+VI is based on the ComE algorithm presented by \citeauthor{ComE} in their paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs} as algorithm 1 on page 381 of the publication's in-proceeding \cite{ComE}. ComE BGMM+VI takes a Bayesian approach to community embedding by utilizing  a Bayesian Gaussian mixture model (BGMM) and variational inference (VI) for community embedding and detection instead of a non-Bayesian Gaussian mixture model (GMM) with expectation maximization (EM) \cite{ComE_BGMM_GH}. %TODO citation to the ComE BGMM paper.

The runtime of ComE BGMM+VI is equivalent to the runtime of ComE in its big-$O$ notation. ComE's time complexity is $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K))$, which is linear in time complexity to the graph size: $O(\abs{V} + \abs{E})$ (line 1) \cite{ComE}. All friends of each user are determined in $O(E)$ (lines 2-5). A sorted dictionary of the top-$N$ friend suggestions for each user sorted ascending by distances between the node embeddings is computed in $O(\abs{V}^2N))$ (lines 6-13). Top-$N$ friend suggestions are generated for each user: $O(\abs{V})$. For each user, all users not befriended currently are considered ($O(\abs{V})$) and inserted into a sorted dictionary of length $N$ ($O(N)$).

This brings the total time complexity of Algorithm \ref{alg:friend_sug} to $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K) + \abs{E} + \abs{V}^2N)$. We consider $\gamma$, $\ell$, $T_1$, $T_2$, and $N$ as constant, therefore the time complexity depends on the graph's size $(\abs{V}, \abs{E})$ and the number of communities $K$: $O(K + \abs{V}K + \abs{E} + \abs{V}^2)$.

Algorithm \ref{alg:friend_sug_com} describes in pseudocode how recommendations are computed from ComE BGMM+VI node and community embeddings and community assignments. The advantage of also considering community membership, is that an improvement in time complexity can be obtained, by considering only users in the same community for friend suggestions.

\begin{algorithm}
    \caption{Top-$N$ Social Recommendations based on Node and Community Embeddings}
    \label{alg:friend_sug_com}
    \begin{algorithmic}[1]
        \Require{graph $G = (V, E)$, maximum number of communities K, number of walks $\gamma$, walk length $\ell$, window size $\zeta$, representation size $D$, negative context size $m$, parameters ($\alpha$, $\beta$), number of recommendations $N$.}
        \Ensure{Top-$N$ recommendations for all nodes $R$.}
        % run ComE BGMM+VI
        \State $\Phi, \Phi', \Pi, (\Psi,\Sigma) \leftarrow ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$
        %  generate a set of nodes for each community
        \For{$v \in V$}
        \State $k \leftarrow \pi_v$
        \State $C_k \leftarrow C_k \cup \{v\}$
        \EndFor
        % find friends
        \For{$(v, v') \in E$}
        \State $F_v \leftarrow F_v \cup \{v'\}$
        \State $F_{v'} \leftarrow F_{v'} \cup \{v\}$
        \EndFor
        % for each node
        \For{$v \in V$}
        % generate ascending recommendation list for each node
        \State $R_v \leftarrow SortedDict(size=N)$
        \For{$v' \in C_k \land v' \notin F_v \land v' \neq v$}
        \State $d \leftarrow \|\phi_v - \phi_{v'}\|$
        \State $R_v[d] \leftarrow v'$
        \EndFor
        % get Top-N recommendations for each node
        \State $R_v  \leftarrow R_v.values()$
        \EndFor
    \end{algorithmic}
\end{algorithm}

Line 1, lines 6-9, and lines 10-17, except for line 12, are the same in Algorithm \ref{alg:friend_sug_com} as in Algorithm \ref{alg:friend_sug}. All users are filtered into sets $C_k$, one for each community $k$ in $O(\abs{V})$ (lines 2-5). Instead of considering the set of all non-friends users of size $\abs{V}$, only users in the same community $C_k$ of size $\frac{\abs{V}}{K}$ are considered (line 12) in $O(\frac{\abs{V}^2N}{K})$.

This brings the total time complexity of Algorithm \ref{alg:friend_sug_com} to $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K) + \abs{V} + \abs{E} + \frac{\abs{V}^2N}{K})$. Again, we consider $\gamma$, $\ell$, $T_1$, $T_2$, and $N$ as constant, therefore the time complexity depends on the graph's size $(\abs{V}, \abs{E})$ and the number of communities $K$: $O(K + \abs{V}K + \abs{E} + \frac{\abs{V}^2}{K})$.

The runtimes for Algorithm \ref{alg:friend_sug} and \ref{alg:friend_sug_com} when considering $K$ as constant or scaling w.r.t. the graph $G = (V, E)$ are shown in Table \ref{tab:runtimes}. The number of friend suggestions $N$ is considered constant w.r.t. the graph $G = (V, E)$.

\begin{table}
    \centering
    \caption{Time complexity of algorithms \ref{alg:friend_sug} and \ref{alg:friend_sug_com}.}
    \label{tab:runtimes}
    \begin{tabular}{l l | c}
        \multirow{2}{*}{$K$ constant} & Algorithm \ref{alg:friend_sug}     & $O(\abs{E} + \abs{V}^2)$                          \\
                                      & Algorithm \ref{alg:friend_sug_com} & $O(\abs{E} + \abs{V}^2)$                          \\
        \hline
        \multirow{2}{*}{$K$ scaling}  & Algorithm \ref{alg:friend_sug}     & $O(K + \abs{V}K + \abs{E} + \abs{V}^2)$           \\
                                      & Algorithm \ref{alg:friend_sug_com} & $O(K + \abs{V}K + \abs{E} + \frac{\abs{V}^2}{K})$ \\
    \end{tabular}
\end{table}

When the hyperparameter number of communities $K$ is considered scaling w.r.t. the graph $G$'s size $(\abs{V}, \abs{E})$, Algorithm \ref{alg:friend_sug_com} reduces the quadratic runtime in comparison to Algorithm \ref{alg:friend_sug}.
$K$ can therefore reduce the quadratic summand $\abs{V}^2N$ by only considering users of the same community for friend suggestions. The reduction in time complexity through Algorithm \ref{alg:friend_sug_com} is only scalable when considering $K$ to be scale with the input graph's size. This assumption can be made when considering communities as groups of friends: with ten times more users, these users form ten times more communities.

\section{Visual Example}

The small \textit{Zachary’s Karate Club graph} dataset will be used to visually underline the motivation for using node embedding and community memberships generated by ComE BGMM+VI for generating friend suggestions for person nodes on the Karate Club graph \cite{KarateClubDS}.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.4\textwidth]{images/karateclub/spring.png}
    \caption{\textit{Zachary’s Karate Club graph} plotted with networkx's spring layout \cite{KarateClubDS, networkx}.}
    \label{fig:karateclub_spring}
\end{figure}

Fig. \ref{fig:karateclub_spring} shows the 34 nodes and 78 edges of the \textit{Zachary’s Karate Club graph} dataset plotted as a graph with the Python network package networkx's spring layout \cite{networkx}. Spring layout simulates a force-directed representation of the network, by treating edges as springs pulling nodes together and treating nodes as repelling objects, sometimes called an anti-gravity force \cite{networkx}. The node classification, represented by node color, shown in Fig. \ref{fig:karateclub_spring} was obtained by running ComE with BGMM and VI with the hyperparameters presented in Table \ref{tab:params_karateclub}.

\begin{table}
    \centering
    \caption{Hyperparameters used for the Karateclub dataset.}
    \label{tab:params_karateclub}
    \begin{tabular}{ l | c | r }
        parameter                    & notation & value     \\
        \hline
        \hline
        number\_walks                & $\gamma$ & $10$      \\
        \hline
        walk\_length                 & $\ell$   & $80$      \\
        \hline
        representation\_size         & $D$      & $2$       \\
        \hline
        num\_workers                 & \empty   & $10$      \\
        \hline
        num\_iter                    & \empty   & $3$       \\
        \hline
        reg\_covar                   & \empty   & $0.00001$ \\
        \hline
        batch\_size                  & \empty   & $50$      \\
        \hline
        window\_size                 & $\zeta$  & $10$      \\
        \hline
        negative                     & $m$      & $5$       \\
        \hline
        lr                           & \empty   & $0.025$   \\
        \hline
        alpha                        & $\alpha$ & $0.1$     \\
        \hline
        beta                         & $\beta$  & $0.1$     \\
        \hline
        down\_sampling               & \empty   & $0.0$     \\
        \hline
        communities                  & $K$      & $5$       \\
        \hline
        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$ \\
    \end{tabular}
\end{table}

As can be seen in Fig. \ref{fig:karateclub_spring}, the two communities in the \textit{Zachary’s Karate Club graph} dataset are correctly identified, despite the number of communities being initialized with $K=5$. Thanks to the Bayesian approach to community modeling and optimization through variational inference, three unused communities are dropped, which leaves the prominent two communities present in the dataset. The node classification obtained reflects the node classification published in the \citeyear{ComE} original ComE paper by \citeauthor{ComE} \cite{ComE}.

\subsection{Training and Testing Split}

\textit{Zachary’s Karate Club graph} is split into training and testing parts. For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.

The three users with the highest degree are chosen as testing users: user 30, user 0, and user 32. For each testing user, a LOO-testing edge is chosen: (30, 23), (0, 1), and (32, 30). These three testing edges are spared for the testing dataset and omitted from the training graph.

\subsection{Embedding}

ComE BGMM+VI is run with the hyperparameters listed in Table \ref{tab:params_karateclub} on the training dataset. The resulting node and community embeddings are visualized as node embeddings colored by community membership and community embeddings plotted as ellipses in Fig. \ref{fig:karateclub_emb}.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=1\linewidth]{images/karateclub/embeddings_5.png}
    \caption{\textit{Zachary’s Karate Club graph} training dataset node and community embedding and node classification.}
    \label{fig:karateclub_emb}
\end{figure}

\subsection{Friend Suggesting}

With embeddings generated, top-$N$ friend suggestion recommendations are generated from the training node and community embeddings for the testing users. For each testing user, a ranked list of non-friend users from the same community is generated (Algorithm \ref{alg:friend_sug_com}). A test user's friend suggestion list is ordered in ascending order by the Euclidean distance between the node embeddings. Table \ref{tab:karateclub_emb_dists} shows the test users' friend suggestion ranked lists.

\begin{table}
    \centering
    \caption{Testing users' friend suggestions ranked by node embedding distance.}
    \label{tab:karateclub_emb_dists}
    \begin{subfigure}{.4\textwidth}
        \centering
        \begin{tabular}{ c | r }
            User        & Distance   \\
            \hline
            \hline
            24          & $0.137181$ \\
            \hline
            25          & $0.214332$ \\
            \hline
            \textbf{23} & $0.543287$ \\
        \end{tabular}
        \caption{user 33}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \begin{tabular}{ c | r }
            User       & Distance   \\
            \hline
            \hline
            \textbf{1} & $0.322270$ \\
            \hline
            16         & $1.030336$ \\
        \end{tabular}
        \caption{user 0}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \begin{tabular}{ c | r }
            User        & Distance   \\
            \hline
            \hline
            25          & $0.039816$ \\
            \hline
            24          & $0.045994$ \\
            \hline
            27          & $0.182484$ \\
            \hline
            26          & $0.255288$ \\
            \hline
            28          & $0.356683$ \\
            \hline
            9           & $0.501317$ \\
            \hline
            \textbf{30} & $0.668306$ \\
        \end{tabular}
        \caption{user 32}
    \end{subfigure}
\end{table}

The testing users' removed edges make it to the lists successfully; therefore, the removed friends are in the same community as the test users and can be recommended through friend suggestions.

\subsection{Evaluation}

The top-$N$ friend suggestions, generated for the testing users with Algorithm \ref{alg:friend_sug_com}, are evaluated against the three removed test edges.
The ranked lists of friend-suggestions shown in Table \ref{tab:karateclub_emb_dists} are used to determine top-$N$ friend suggestions for each testing user and evaluate the recommendations by the hit-rate metric. The LOO-testing edges are printed bold in Table \ref{tab:karateclub_emb_dists}. The resulting top-$N$ hit-rates are plotted against the number of suggestions $N$ in Fig. \ref{fig:karateclub_hitrates}.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=.4\textwidth]{images/karateclub/hit_rates.png}
    \caption{Hit-rates from top-$N$ recommendations on \textit{Zachary’s Karate Club graph}.}
    \label{fig:karateclub_hitrates}
\end{figure}

Fig. \ref{fig:karateclub_hitrates} shows promising results for using ComE to generate top-$N$ friend suggestions on the small \textit{Zachary’s Karate Club graph}. The achieved hit-rate starts at 33\% for $N=1$, reaches 67\% at $N=3$, and maxes out at 100\% for $N=7$; therefore, when presenting three friend suggestions to one test user, at least in $66\%$ of cases a friend suggestion would be of interest.

\section{Facebook Friend Suggestions}

The recommender system based on community embeddings with the Bayesian Gaussian mixture model and variational inference is quantitatively evaluated. Friend suggestions are generated for the \textit{Social Circles: Facebook} dataset \cite{social_circles_fb, social_circles_in_ego_networks}.

The method of generating friend suggestions using ComE BGMM+VI embeddings and the smallest-distance approach is applied to the \textit{Social Circles: Facebook} graph. Both Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} are evaluated. The top-$N$ recommendations are evaluated by the hit-rate metric.

\subsection{Dataset}

The \textit{Social Circles: Facebook} dataset was sourced from the profile and network data of ten ego-networks (egonets) through a custom Facebook app by \citeauthor{social_circles_in_ego_networks} as part of their \citeyear{social_circles_in_ego_networks} article \textit{Learning to Discover Social Circles in Ego Networks} \cite{social_circles_in_ego_networks}.

Fig. \ref{fig:egonet} shows an illustration of the ego-network of user $u$. User $u$ has direct connections to $v_i$, called alters. Alters can be connected to each other and build circles of users, like families, high school friends, or college friends.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.48\textwidth]{images/fb/egonet.png}
    \caption{Illustrative example of an ego-network \cite{social_circles_in_ego_networks}.}
    \label{fig:egonet}
\end{figure}

\citeauthor{social_circles_in_ego_networks} obtained profile and network data from ten Facebook ego-networks. The dataset contains 4039 users (nodes) and 88234 friendships (edges).

\subsection{Training and Testing Split}

The \textit{Social Circles: Facebook} graph is split into training and testing parts. For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.
Users with only one connection are removed from the dataset since no network information is added by them.

$1169$ of the 4039 users of the \textit{Social Circles: Facebook} graph have at least $50$ connections and are used for testing friend suggestions. The threshold of $50$ friends for considering users for testing is set to make sure enough data is available to make sensable suggestions for testing users.

Now that the testing users are identified, the leave-one-out method is employed for obtaining one testing friendship per testing user. The friendship chosen to leave out for testing for each testing user is chosen randomly using NumPy's random choice \cite{oliphant2006guide, van2011numpy}.

\subsection{Embedding}

Community embeddings for the \textit{Social Circles: Facebook} graph are generated by running ComE BGMM+VI with the hyperparameters listed in Table \ref{tab:params_fb}.

\begin{table}
    \centering
    \caption{Hyperparameters used for the \textit{Social Circles: Facebook} dataset.}
    \label{tab:params_fb}
    \begin{tabular}{ l | c | r }
        parameter                    & notation & value       \\
        \hline
        \hline
        number\_walks                & $\gamma$ & $10$        \\
        \hline
        walk\_length                 & $\ell$   & $80$        \\
        \hline
        representation\_size         & $D$      & $2, 4, 128$ \\
        \hline
        num\_workers                 & \empty   & $10$        \\
        \hline
        num\_iter                    & \empty   & $3$         \\
        \hline
        com\_n\_init                 & \empty   & $10$        \\
        \hline
        reg\_covar                   & \empty   & $0.00001$   \\
        \hline
        batch\_size                  & \empty   & $50$        \\
        \hline
        window\_size                 & $\zeta$  & $10$        \\
        \hline
        negative                     & $m$      & $5$         \\
        \hline
        lr                           & \empty   & $0.025$     \\
        \hline
        alpha                        & $\alpha$ & $0.1$       \\
        \hline
        beta                         & $\beta$  & $0.1$       \\
        \hline
        down\_sampling               & \empty   & $0.0$       \\
        \hline
        communities                  & $K$      & $10$        \\
        \hline
        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$   \\
    \end{tabular}
\end{table}

As is shown in Table \ref{tab:params_fb}, the embedding's representation size $D$ is varied along three values. A representation size $D \in [2,4,128]$ is chosen to determine the effect of the embedding's dimensionality on friend suggestions. It is hypothesized, that two dimensions are not sufficient to capture the structure of the \textit{Social Circles: Facebook} graph, due to its size, therefore higher dimensional embeddings are evaluated as well.

Fig. \ref{fig:fb_emb_anim_d2}, Fig. \ref{fig:fb_emb_anim_d4}, and Fig. \ref{fig:fb_emb_anim_d128} show the node embeddings generated by ComE BGMM+VI after one, two, and three iterations. The nodes are colored by community membership from ComE's community embedding and detection.
The $4$- and $128$-dimensional node and community embeddings cannot effectively be plotted in two-dimensional print. Instead, in Fig. \ref{fig:fb_emb_anim_d4} and Fig. \ref{fig:fb_emb_anim_d128} the two first dimensions of the node and community embeddings are plotted to gain a narrow perspective on the resulting node embeddings.

\newcommand{\fbEmbAnimD}[1]{
    \begin{figure*}[htbp]
        \centering
        \begin{subfigure}{.33\textwidth}
            \centering
            \includegraphics[width=1\linewidth]{images/fb/anim_d#1_i0.png}
            \caption{\nth{1} Iteration}
        \end{subfigure}%
        %
        \begin{subfigure}{.33\textwidth}
            \centering
            \includegraphics[width=1\linewidth]{images/fb/anim_d#1_i1.png}
            \caption{\nth{2} Iteration}
        \end{subfigure}%
        %
        \begin{subfigure}{.33\textwidth}
            \centering
            \includegraphics[width=1\linewidth]{images/fb/anim_d#1_i2.png}
            \caption{\nth{3} Iteration}
        \end{subfigure}%
        %
        \caption{Node embeddings segment with representation size $D=#1$.}
        \label{fig:fb_emb_anim_d#1}
    \end{figure*}
}

\fbEmbAnimD{2}
\fbEmbAnimD{4}
\fbEmbAnimD{128}

Fig. \ref{fig:fb_graphs} shows the graph spring plotted with networkx and nodes colored by community membership after the last iteration of ComE \cite{networkx}. The spring plotted graph structure on Fig. \ref{fig:fb_graphs} suggests, that with embedding dimension $D=4$, more fitting communities are detected than for $D=2$ and $D=128$.

\begin{figure*}[htbp]
    \centering
    \begin{subfigure}{.33\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/graph_d2.png}
        \caption{$D=2$}
    \end{subfigure}%
    %
    \begin{subfigure}{.33\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/graph_d4.png}
        \caption{$D=4$}
    \end{subfigure}%
    %
    \begin{subfigure}{.33\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/graph_d128.png}
        \caption{$D=128$}
    \end{subfigure}%
    %
    \caption{Spring plotted graphs with nodes colored by community membership after last ComE BGMM+VI iteration.}
    \label{fig:fb_graphs}
\end{figure*}

Fig. \ref{fig:fb_emb_anim_d2} shows node and community embeddings with representation size $D=2$. The limiting two-dimensional space does not allow ComE to represent the \textit{Social Circles: Facebook} graph's structure sufficiently to utilize the $K=10$ components of the Gaussian mixture model. Only three components are used after the last of three iterations.

While Fig. \ref{fig:fb_emb_anim_d4} and Fig. \ref{fig:fb_emb_anim_d128} show only the first two dimensions, of the node embedding of 4 or 128 dimensionality respectively, the general structure of communities is visible. The development of the node embeddings over the three iterations shows the manifestation of the first-, second-, and higher-order proximity by ComE's closed loop. The node embeddings are densified in their communities and spread out in regards to other communities. All $K=10$ components of the Gaussian mixture model are used for $D=4$ and $D=128$.

The nodes are distributed over the $K=10$ communities as shown in the histograms in Fig. \ref{fig:fb_emb_weights}. Fig. \ref{fig:fb_emb_weights_d2} shows that with representation size $D=2$, ComE BGMM+VI drops 6 communities, effecitvelly using only 4 communities in the final embedding, because the two-dimensional node embeddings are not expressive enough to support more communities. Fig. \ref{fig:fb_emb_weights_d4} and Fig. \ref{fig:fb_emb_weights_d128} show that with representation sizes $D=4$ and $D=128$ respectivelly, all $K=10$ communities are used.

The tradeoff between the representation size of the embedding $D$ and the embedding expressiveness is demonstrated. When choosing a too low-dimensional embedding, the embedding might not be expressive enough to capture the communities and reduces the quality of the community embedding and detection. When choosing a too high-dimensional embedding, the node embeddings get so expressive that BGMM+VI is not able to drop unused communities because all communities are used in one of the many dimensions.

\begin{figure*}[htbp]
    \centering
    %
    \begin{subfigure}{.33\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/weights_d2.png}
        \caption{$D=2$}
        \label{fig:fb_emb_weights_d2}
    \end{subfigure}%
    %
    \begin{subfigure}{.33\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/weights_d4.png}
        \caption{$D=4$}
        \label{fig:fb_emb_weights_d4}
    \end{subfigure}%
    %
    \begin{subfigure}{.33\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/weights_d128.png}
        \caption{$D=128$}
        \label{fig:fb_emb_weights_d128}
    \end{subfigure}%
    %
    \caption{Weights of communities with $K=10$ and $D \in [2,4,128]$.}
    \label{fig:fb_emb_weights}
\end{figure*}

\subsection{Friend Suggestions}

With node and community embeddings and community assignments generated for $D \in [2,4,128]$ and $K=10$ with training data in the previous section, top-$N$ friend suggestion recommendations are generated from the training node and community embeddings with Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com}.

The shortest-embeddings-distance approach is employed, recommending friends to users who have the shortest distance node embeddings over all users (Algorithm \ref{alg:friend_sug}) or only over users in the same community (Algorithm \ref{alg:friend_sug_com}). Users who are already connected are not recommended again.

\section{Evaluation}

The charts in Fig. \ref{fig:fb_hit_rates} show the hit-rates for top-$N$ recommendations for testing users.

\begin{figure}[htbp]
    \centering
    % d2
    \begin{subfigure}{.48\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/hit_rates_d-2.png}
        \caption{$D=2$ on linear $N$-axis.}
        \label{fig:fb_hit_rates_d2}
    \end{subfigure}%

    % d4
    \begin{subfigure}{.48\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/hit_rates_d-4.png}
        \caption{$D=4$ on linear $N$-axis.}
        \label{fig:fb_hit_rates_d4}
    \end{subfigure}%

    % d128
    \begin{subfigure}{.48\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{images/fb/hit_rates_d-128.png}
        \caption{$D=128$ on linear $N$-axis.}
        \label{fig:fb_hit_rates_d128}
    \end{subfigure}%

    \caption{Hit-rates of \textit{Social Circles: Facebook} graph's top-$N$ friend suggestions on $N$-axis for $D \in [2,4,128]$. Hit-rate lines for Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} overlap for $D=2$ and $D=4$.}
    \label{fig:fb_hit_rates}
\end{figure}

Two hit-rate lines are shown for each embedding representation size $D \in [2,4,128]$ on $N$-axis. The blue line shows the hit-rates when generating friend suggestions from all users (Algorithm \ref{alg:friend_sug}). The orange line shows the hit-rate when generating recommendations from users in the same community (Algorithm \ref{alg:friend_sug_com}). For lower-dimensional embeddings with $D \in [2,4]$, the hit-rates for Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} largely overlap.

Table \ref{tab:fb_hit_rates} shows readings of hit-rates over the algorithms and embedding dimensionality for top-$5$, top-$10$, and top-$20$ recommendations:

\begin{figure}[htbp]
    \centering
    \caption{Hit-rates of \textit{Social Circles: Facebook} graph's top-$N$ friend suggestions for algorithms \ref{alg:friend_sug} and \ref{alg:friend_sug_com}, $N \in [5, 10, 20]$, and $D \in [2, 4, 128]$.}
    \label{tab:fb_hit_rates}
    \begin{tabular}{ l | l | c | c | c }
        \hline
        \empty                                              & \empty  & top-$5$         & top-$10$        & top-$20$        \\
        \hline
        \hline
        \multirow{3}{*}{Algorithm \ref{alg:friend_sug}}     & $D=2$   & 8.5\%           & 14.2\%          & 21.9\%          \\
                                                            & $D=4$   & 18.1\%          & 26.5\%          & 38.4\%          \\
                                                            & $D=128$ & \textbf{31.3\%} & \textbf{43.6\%} & \textbf{58.8\%} \\
        \hline
        \multirow{3}{*}{Algorithm \ref{alg:friend_sug_com}} & $D=2$   & 8.4\%           & 14.1\%          & 21.8\%          \\
                                                            & $D=4$   & 18.0\%          & 26.4\%          & 38.2\%          \\
                                                            & $D=128$ & \textbf{24.1\%} & \textbf{32.9\%} & \textbf{41.8\%} \\
        \hline
    \end{tabular}
\end{figure}

Hit-rates based on top-$5$, top-$10$, and top-$20$ recommendations are chosen since higher $N$ are not practical for recommendations.

Presenting only the top-$10$ friend suggestions to users, a hit-rate of $43.6\%$ when considering all users, or $32.9\%$, when considering only users from the same community, is achieved. A top-$5$ hit-rate of $31.3\%$ (all users) and $24.1\%$ (same community) is achieved, which can be considered viable for business use-cases, especially for only showing 5 recommendations to users. A top-$20$ hit-rate of $58.8\%$ (all users) and $41.8\%$ (same community) adds to the presented results.

\section{Conclusion}

The proposed shortest-embedding-distance friend suggestion recommender system based on embeddings generated with the community embeddings algorithm ComE BGMM+VI shows promising results.
A top-$10$ recommendations hit-rate of $43.6\%$ when considering all users (Algorithm \ref{alg:friend_sug}), and over $32.9\%$ when considering users in the same community (Algorithm \ref{alg:friend_sug_com}) is achieved for Facebook friend suggestions. Both results were achieved with a representation size $D=128$. Lower hit-rates are obtained for $D=2$ and $D=4$. See Table \ref{tab:fb_hit_rates}.

Algorithm \ref{alg:friend_sug} achieves the better hit-rate in comparison to Algorithm \ref{alg:friend_sug_com}, since all users are considered for embedding instead of just users of the same community. Algorithm \ref{alg:friend_sug_com}, on the other hand, has better time complexity than Algorithm \ref{alg:friend_sug}, since only users in the same community are considered for friend suggestions instead of all users. Algorithm \ref{alg:friend_sug_com} is scalable to larger datasets if the number of communities $K$ can be considered to scale with the graph size $(\abs{V}, \abs{E})$.

\subsection{Future Work}

This work explores generating friend suggestions from node and community embeddings generated by ComE. Three values of the representation size $D$ were used in experimentation. Experimenting with more, different combinations of hyperparameters, and especially varying the number of communities $K$ might be of interest. Since in this paper, friend suggestions were explored, it is left for future work to explore item-based recommendations using embeddings. More community-centric recommendations, which deeply consider the community embeddings in generating recommendations, are left for future work.

%%% Bibliography
\bibliographystyle{IEEEtranN}
\bibliography{literature}

\end{document}

### 2020-10-14 paper-ffs.tex
title, abstract and keywords

@@
 
-\title{Community Embeddings for Facebook Friend Suggestions\\
+\title{Community Embeddings with Bayesian Gaussian Mixture Model and Variational Inference\\
     \thanks{The reported study was partially supported by RFBR grant \textnumero 20-07-00958. The paper was prepared within the framework of the HSE University Project Group Competition 2020-2022.}
@@
     Graphs, such as social networks, emerge naturally from various real-world situations. Recently, graph embedding methods have gained traction in data science research.
-    Recommender systems are used in a wide range of business applications and are essential for online e-business models to survive and strive in the contemporary market.
-Using graph embeddings for recommendation tasks can improve upon recommender systems because of data compression, the feature vector format, and sub-quadratic time complexity on computing embeddings.
-    Graph and community embeddings generated with ComE BGMM+VI are used to build a recommender system for friend suggestions.
-ComE BGMM+VI is an alteration of the community embeddings algorithm ComE.
-ComE BGMM+VI applies a Bayesian Gaussian mixture model and variational inference for community embedding and detection.
-    Recommendations are evaluated by the top-N hit-rate for users with at least 50 friends.
-A friend suggestions recommender system with a top-10 leave-one-out hit-rate of 43.6\% and run-time optimized hit-rate of 32.9\% is presented.
+    The graph and community embedding algorithm ComE aims to preserve first-, second- and higher-order proximity.
+ComE requires prior knowledge of the number of communities K.
+In this paper, ComE is extended to utilize a Bayesian Gaussian mixture model with variational inference for learning community embeddings (ComE BGMM+VI), similar to ComE+.
+ComE BGMM+VI takes K as the maximum number of communities and drops components through the trade-off hyperparameter weight concentration prior.
+    The advantage of ComE BGMM+VI over the non-Bayesian ComE for an unknown number of communities K is shown for the small Karate club dataset and explored for the larger DBLP dataset.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-    graph, embedding, community embedding, ComE, recommender systems, friend suggestions
+    graph, embedding, community embedding, ComE, Bayesian, variational inference, Gaussian mixture, expectation maximization
 \end{IEEEkeywords}

### 2020-10-14 paper-ffs.tex
remove FSs content

@@
 
-Graphs, such as social networks, knowledge graphs, content-rating graphs, and communication networks emerge naturally from various real-world situations.
-Analyzing these graphs leads to findings and understanding of the underlying structures, coherences, and dependencies.
-Recently, methods for embedding graph's nodes into lower-dimensional, Euclidean spaces, called graph embeddings, have gained traction in multiple areas of data science research \cite{Goyal_2018}.
-
-Due to the rapid growth of the internet and data accumulation, recommender systems are essential for e-business and online business models to survive and strive in the contemporary market \cite{Polatidis2013RecommenderST}.
-Modern recommender systems need to take into account the huge amounts of data generated at all times in big data systems around the world and improve recommendations instead of failing under the thrust of big data overload.
-
-Utilizing graph embeddings for recommendation tasks, has recently gained research interest \cite{Palumbo2018AnEC, Palumbo2018KnowledgeGE, GradGyenge2017GraphEB, Sathish2019GraphEB}.
-The advantages of graph embeddings include data compression and the feature vector format \cite{Godec2018}.
-Given these advantages and provided competitive results, graph embeddings have the possibility of greatly improving upon graph-based use-cases like recommender systems.
-
-Community Embeddings, in addition to embedding a graph's nodes through first- and second-order proximity, also preserve higher-order proximity by embedding clusters present in the graph data.
-The graph and community embedding algorithm ComE aims to preserve first-, second- and higher-order proximity by embedding a graph's nodes and communities \cite{ComE}.
-
-This work specifically examines community embeddings for friend suggestion recommender systems and evaluates recommendations on social network graph data for the use-case of friend suggestions.
-Graph and community embeddings generated with ComE BGMM+VI are used to develop a friend suggestions recommender system based on the shortest distances between nodes in the embedding.
-Recommendations are evaluated by the top-$N$ recommendations hit-rate of test edges.
-A friend suggestions recommender system with a top-10 leave-one-out hit-rate of 43.6\% and run-time optimized 32.9\% is presented.
-
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width=0.48\textwidth]{images/fb/graph_d4.png}
-    \caption{\textit{Social Circles: Facebook} graph plotted with networkx's spring layout and node colors representing community membership obtained by ComE BGMM+VI with $D=4$ \cite{networkx}.}
-    \label{fig:fb_graph}
-\end{figure}
-
-\section{Friend Suggestions}
-
-Recommender Systems are eagerly researched in academia and widely deployed in real-world business applications.
-Most contemporary technology companies heavily rely on recommender systems to drive usage of their services and consumption of their content.
-Users, in turn, rely on recommender systems to find what they want, need, save searching time, and what they might not have found otherwise.
-State-of-the-art recommender systems provide a competitive advantage desperately needed by online services.
-Companies heavily relying on recommender systems include YouTube, Amazon, Netflix, and many more \cite{Rocca19}.
-
-Recommender systems are built on top of user-item interaction.
-Friend suggestions are a simpler type of recommender system.
-For friend suggestions no distinction is made between users and items.
-The entity user is both the subject and object of recommendation.
-This results in a simple data model, which can be used for structural friend suggestions, as shown in Fig.
-\ref{fig:data_model}.
-
-\begin{figure}[htbp]
-    \centering
-    \begin{tikzpicture}[node distance=6cm]
-        \node (user) [user] {User};
-        \path [arrow] (user) edge[loop right] node[right]{relate} (user);
-    \end{tikzpicture}
-    \caption{Minimal data model supporting friend suggestions.}
-    \label{fig:data_model}
-\end{figure}
-
-The two main methods used for recommender systems are collaborative filtering and content-based filtering.
-Collaborative filtering is based on the assumption that users like items similar to other items they like and items that are liked by other users with similar tastes \cite{Grover17, Su2009ASO}.
-Content-based filtering considers user and item attributes, instead of solely interactions \cite{Rocca19, Lops2011SurveyCB}.
-
-In this paper, a method of generating friend suggestions based solely on graph data structure, and not on the node or edge attributes, is presented.
-The graph of users and friendships is embedded to a lower-dimensional space with the community embedding algorithm ComE BGMM+VI \cite{ComE, ComE+, ComE_BGMM_GH}.
-The approach can be considered a variant of collaborative filtering with the Euclidean distance of user embeddings and user community membership as measures for user similarity.
-The proposed advantage of using such an approach is that community embeddings optimize first-, second-, and higher-order proximity between users in the node embedding.
-
-\subsection{Evaluation}
-
-To determine the effectivity of friend suggestions, the generated friend suggestions must be properly validated.
-In this paper, the top-$N$ approach to evaluating recommender systems with hit-rate as the evaluation metric is chosen.
-
-A user's top-$N$ recommendations is a list of $N$ items to be recommended to a specific user.
-To evaluate recommender systems with the top-$N$ and hit-rate approach, initially, the dataset is split into train and test data.
-For each testing user, one relation is left out, according to the leave-one-out method and the model is trained on the remaining training dataset.
-Once the model is trained, a list of top-$N$ recommendations is generated for each testing user.
-If the item corresponding to the user is in the user's top-$N$ list, a hit is counted, otherwise, a miss is counted.
-The hit-rate is defined as the total number of hits divided by the number of testing users.
-
-Utilizing the top-$N$ approach with hit-rate to evaluate recommender systems is advantageous to evaluating recommender systems by a link prediction approach since the top-$N$ approach is more realistic in comparison to actual recommender system use-cases.
-When you open Netflix, Amazon, YouTube, or Facebook one or multiple top-$N$ recommendation lists are generated and displayed.
-If you click on an item and buy, watch, or befriend, that is considered a hit, otherwise, a miss.
-
-The hit-rate metric on top-$N$ recommendations provides a realistic option of evaluating recommender systems and is often used with the leave-one-out (LOO) strategy in literature \cite{Cremonesi2010PerformanceOR, Palumbo2017entity2rec, Zhao2019ATS, He2017NeuralCF}.
-
-\section{Algorithm}
-
-The proposed algorithm for generating friend suggestions using community embeddings is detailed and evaluated on time complexity.
-The initially quadratic runtime of generating friend suggestions for all users is then reduced by a factor of $K$ by utilizing a node's community membership.
-The two resulting algorithms are presented.
-
-Algorithm \ref{alg:friend_sug} describes in pseudocode how recommendations are computed from node embeddings generated by ComE BGMM+VI.
-The terms friend suggestions and social recommendations are used interchangeably.
-
-\begin{algorithm}
-    \caption{Top-$N$ Social Recommendations based on Node Embeddings}
-    \label{alg:friend_sug}
-    \begin{algorithmic}[1]
-        \Require{graph $G = (V, E)$, maximum number of communities K, number of walks $\gamma$, walk length $\ell$, window size $\zeta$, representation size $D$, negative context size $m$, parameters ($\alpha$, $\beta$), number of recommendations $N$.}
-        \Ensure{Top-$N$ recommendations for all nodes $R$.}
-        % run ComE BGMM+VI
-        \State $\Phi, \Phi', \Pi, (\Psi,\Sigma) \leftarrow ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$
-        % find friends
-        \For{$(v, v') \in E$}
-        \State $F_v \leftarrow F_v \cup \{v'\}$
-        \State $F_{v'} \leftarrow F_{v'} \cup \{v\}$
-        \EndFor
-        % for each node
-        \For{$v \in V$}
-        % generate ascending recommendation list for each node
-        \State $R_v \leftarrow SortedDict(size=N)$
-        \For{$v' \in V \land v' \notin F_v \land v' \neq v$}
-        \State $d \leftarrow \|\phi_v - \phi_{v'}\|$
-        \State $R_v[d] \leftarrow v'$
-        \EndFor
-        % get Top-N recommendations for each node
-        \State $R_v  \leftarrow R_v.values()$
-        \EndFor
-    \end{algorithmic}
-\end{algorithm}
-
-The function $ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ is the ComE BGMM+VI community embedding algorithm.
-It returns the node embedding $\Phi$, context embedding $\Phi'$, community assignment $\Pi$, and community embedding $(\Psi,\Sigma)$.
-ComE BGMM+VI is based on the ComE algorithm presented by \citeauthor{ComE} in their paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs} as algorithm 1 on page 381 of the publication's in-proceeding \cite{ComE}.
-ComE BGMM+VI takes a Bayesian approach to community embedding by utilizing  a Bayesian Gaussian mixture model (BGMM) and variational inference (VI) for community embedding and detection instead of a non-Bayesian Gaussian mixture model (GMM) with expectation maximization (EM) \cite{ComE_BGMM_GH}.
-%TODO citation to the ComE BGMM paper.
-
-The runtime of ComE BGMM+VI is equivalent to the runtime of ComE in its big-$O$ notation.
-ComE's time complexity is $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K))$, which is linear in time complexity to the graph size: $O(\abs{V} + \abs{E})$ (line 1) \cite{ComE}.
-All friends of each user are determined in $O(E)$ (lines 2-5).
-A sorted dictionary of the top-$N$ friend suggestions for each user sorted ascending by distances between the node embeddings is computed in $O(\abs{V}^2N))$ (lines 6-13).
-Top-$N$ friend suggestions are generated for each user: $O(\abs{V})$.
-For each user, all users not befriended currently are considered ($O(\abs{V})$) and inserted into a sorted dictionary of length $N$ ($O(N)$).
-
-This brings the total time complexity of Algorithm \ref{alg:friend_sug} to $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K) + \abs{E} + \abs{V}^2N)$.
-We consider $\gamma$, $\ell$, $T_1$, $T_2$, and $N$ as constant, therefore the time complexity depends on the graph's size $(\abs{V}, \abs{E})$ and the number of communities $K$: $O(K + \abs{V}K + \abs{E} + \abs{V}^2)$.
-
-Algorithm \ref{alg:friend_sug_com} describes in pseudocode how recommendations are computed from ComE BGMM+VI node and community embeddings and community assignments.
-The advantage of also considering community membership, is that an improvement in time complexity can be obtained, by considering only users in the same community for friend suggestions.
-
-\begin{algorithm}
-    \caption{Top-$N$ Social Recommendations based on Node and Community Embeddings}
-    \label{alg:friend_sug_com}
-    \begin{algorithmic}[1]
-        \Require{graph $G = (V, E)$, maximum number of communities K, number of walks $\gamma$, walk length $\ell$, window size $\zeta$, representation size $D$, negative context size $m$, parameters ($\alpha$, $\beta$), number of recommendations $N$.}
-        \Ensure{Top-$N$ recommendations for all nodes $R$.}
-        % run ComE BGMM+VI
-        \State $\Phi, \Phi', \Pi, (\Psi,\Sigma) \leftarrow ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$
-        %  generate a set of nodes for each community
-        \For{$v \in V$}
-        \State $k \leftarrow \pi_v$
-        \State $C_k \leftarrow C_k \cup \{v\}$
-        \EndFor
-        % find friends
-        \For{$(v, v') \in E$}
-        \State $F_v \leftarrow F_v \cup \{v'\}$
-        \State $F_{v'} \leftarrow F_{v'} \cup \{v\}$
-        \EndFor
-        % for each node
-        \For{$v \in V$}
-        % generate ascending recommendation list for each node
-        \State $R_v \leftarrow SortedDict(size=N)$
-        \For{$v' \in C_k \land v' \notin F_v \land v' \neq v$}
-        \State $d \leftarrow \|\phi_v - \phi_{v'}\|$
-        \State $R_v[d] \leftarrow v'$
-        \EndFor
-        % get Top-N recommendations for each node
-        \State $R_v  \leftarrow R_v.values()$
-        \EndFor
-    \end{algorithmic}
-\end{algorithm}
-
-Line 1, lines 6-9, and lines 10-17, except for line 12, are the same in Algorithm \ref{alg:friend_sug_com} as in Algorithm \ref{alg:friend_sug}.
-All users are filtered into sets $C_k$, one for each community $k$ in $O(\abs{V})$ (lines 2-5).
-Instead of considering the set of all non-friends users of size $\abs{V}$, only users in the same community $C_k$ of size $\frac{\abs{V}}{K}$ are considered (line 12) in $O(\frac{\abs{V}^2N}{K})$.
-
-This brings the total time complexity of Algorithm \ref{alg:friend_sug_com} to $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K) + \abs{V} + \abs{E} + \frac{\abs{V}^2N}{K})$.
-Again, we consider $\gamma$, $\ell$, $T_1$, $T_2$, and $N$ as constant, therefore the time complexity depends on the graph's size $(\abs{V}, \abs{E})$ and the number of communities $K$: $O(K + \abs{V}K + \abs{E} + \frac{\abs{V}^2}{K})$.
-
-The runtimes for Algorithm \ref{alg:friend_sug} and \ref{alg:friend_sug_com} when considering $K$ as constant or scaling w.r.t.
-the graph $G = (V, E)$ are shown in Table \ref{tab:runtimes}.
-The number of friend suggestions $N$ is considered constant w.r.t.
-the graph $G = (V, E)$.
-
-\begin{table}
-    \centering
-    \caption{Time complexity of algorithms \ref{alg:friend_sug} and \ref{alg:friend_sug_com}.}
-    \label{tab:runtimes}
-    \begin{tabular}{l l | c}
-        \multirow{2}{*}{$K$ constant} & Algorithm \ref{alg:friend_sug}     & $O(\abs{E} + \abs{V}^2)$                          \\
-                                      & Algorithm \ref{alg:friend_sug_com} & $O(\abs{E} + \abs{V}^2)$                          \\
-        \hline
-        \multirow{2}{*}{$K$ scaling}  & Algorithm \ref{alg:friend_sug}     & $O(K + \abs{V}K + \abs{E} + \abs{V}^2)$           \\
-                                      & Algorithm \ref{alg:friend_sug_com} & $O(K + \abs{V}K + \abs{E} + \frac{\abs{V}^2}{K})$ \\
-    \end{tabular}
-\end{table}
-
-When the hyperparameter number of communities $K$ is considered scaling w.r.t.
-the graph $G$'s size $(\abs{V}, \abs{E})$, Algorithm \ref{alg:friend_sug_com} reduces the quadratic runtime in comparison to Algorithm \ref{alg:friend_sug}.
-$K$ can therefore reduce the quadratic summand $\abs{V}^2N$ by only considering users of the same community for friend suggestions.
-The reduction in time complexity through Algorithm \ref{alg:friend_sug_com} is only scalable when considering $K$ to be scale with the input graph's size.
-This assumption can be made when considering communities as groups of friends: with ten times more users, these users form ten times more communities.
-
-\section{Visual Example}
-
-The small \textit{Zachary’s Karate Club graph} dataset will be used to visually underline the motivation for using node embedding and community memberships generated by ComE BGMM+VI for generating friend suggestions for person nodes on the Karate Club graph \cite{KarateClubDS}.
-
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width=0.4\textwidth]{images/karateclub/spring.png}
-    \caption{\textit{Zachary’s Karate Club graph} plotted with networkx's spring layout \cite{KarateClubDS, networkx}.}
-    \label{fig:karateclub_spring}
-\end{figure}
-
-Fig.
-\ref{fig:karateclub_spring} shows the 34 nodes and 78 edges of the \textit{Zachary’s Karate Club graph} dataset plotted as a graph with the Python network package networkx's spring layout \cite{networkx}.
-Spring layout simulates a force-directed representation of the network, by treating edges as springs pulling nodes together and treating nodes as repelling objects, sometimes called an anti-gravity force \cite{networkx}.
-The node classification, represented by node color, shown in Fig.
-\ref{fig:karateclub_spring} was obtained by running ComE with BGMM and VI with the hyperparameters presented in Table \ref{tab:params_karateclub}.
-
-\begin{table}
-    \centering
-    \caption{Hyperparameters used for the Karateclub dataset.}
-    \label{tab:params_karateclub}
-    \begin{tabular}{ l | c | r }
-        parameter                    & notation & value     \\
-        \hline
-        \hline
-        number\_walks                & $\gamma$ & $10$      \\
-        \hline
-        walk\_length                 & $\ell$   & $80$      \\
-        \hline
-        representation\_size         & $D$      & $2$       \\
-        \hline
-        num\_workers                 & \empty   & $10$      \\
-        \hline
-        num\_iter                    & \empty   & $3$       \\
-        \hline
-        reg\_covar                   & \empty   & $0.00001$ \\
-        \hline
-        batch\_size                  & \empty   & $50$      \\
-        \hline
-        window\_size                 & $\zeta$  & $10$      \\
-        \hline
-        negative                     & $m$      & $5$       \\
-        \hline
-        lr                           & \empty   & $0.025$   \\
-        \hline
-        alpha                        & $\alpha$ & $0.1$     \\
-        \hline
-        beta                         & $\beta$  & $0.1$     \\
-        \hline
-        down\_sampling               & \empty   & $0.0$     \\
-        \hline
-        communities                  & $K$      & $5$       \\
-        \hline
-        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$ \\
-    \end{tabular}
-\end{table}
-
-As can be seen in Fig.
-\ref{fig:karateclub_spring}, the two communities in the \textit{Zachary’s Karate Club graph} dataset are correctly identified, despite the number of communities being initialized with $K=5$.
-Thanks to the Bayesian approach to community modeling and optimization through variational inference, three unused communities are dropped, which leaves the prominent two communities present in the dataset.
-The node classification obtained reflects the node classification published in the \citeyear{ComE} original ComE paper by \citeauthor{ComE} \cite{ComE}.
-
-\subsection{Training and Testing Split}
-
-\textit{Zachary’s Karate Club graph} is split into training and testing parts.
-For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.
-
-The three users with the highest degree are chosen as testing users: user 30, user 0, and user 32.
-For each testing user, a LOO-testing edge is chosen: (30, 23), (0, 1), and (32, 30).
-These three testing edges are spared for the testing dataset and omitted from the training graph.
-
-\subsection{Embedding}
-
-ComE BGMM+VI is run with the hyperparameters listed in Table \ref{tab:params_karateclub} on the training dataset.
-The resulting node and community embeddings are visualized as node embeddings colored by community membership and community embeddings plotted as ellipses in Fig.
-\ref{fig:karateclub_emb}.
-
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width=1\linewidth]{images/karateclub/embeddings_5.png}
-    \caption{\textit{Zachary’s Karate Club graph} training dataset node and community embedding and node classification.}
-    \label{fig:karateclub_emb}
-\end{figure}
-
-\subsection{Friend Suggesting}
-
-With embeddings generated, top-$N$ friend suggestion recommendations are generated from the training node and community embeddings for the testing users.
-For each testing user, a ranked list of non-friend users from the same community is generated (Algorithm \ref{alg:friend_sug_com}).
-A test user's friend suggestion list is ordered in ascending order by the Euclidean distance between the node embeddings.
-Table \ref{tab:karateclub_emb_dists} shows the test users' friend suggestion ranked lists.
-
-\begin{table}
-    \centering
-    \caption{Testing users' friend suggestions ranked by node embedding distance.}
-    \label{tab:karateclub_emb_dists}
-    \begin{subfigure}{.4\textwidth}
-        \centering
-        \begin{tabular}{ c | r }
-            User        & Distance   \\
-            \hline
-            \hline
-            24          & $0.137181$ \\
-            \hline
-            25          & $0.214332$ \\
-            \hline
-            \textbf{23} & $0.543287$ \\
-        \end{tabular}
-        \caption{user 33}
-    \end{subfigure}
-    \begin{subfigure}{.4\textwidth}
-        \centering
-        \begin{tabular}{ c | r }
-            User       & Distance   \\
-            \hline
-            \hline
-            \textbf{1} & $0.322270$ \\
-            \hline
-            16         & $1.030336$ \\
-        \end{tabular}
-        \caption{user 0}
-    \end{subfigure}
-    \begin{subfigure}{.4\textwidth}
-        \centering
-        \begin{tabular}{ c | r }
-            User        & Distance   \\
-            \hline
-            \hline
-            25          & $0.039816$ \\
-            \hline
-            24          & $0.045994$ \\
-            \hline
-            27          & $0.182484$ \\
-            \hline
-            26          & $0.255288$ \\
-            \hline
-            28          & $0.356683$ \\
-            \hline
-            9           & $0.501317$ \\
-            \hline
-            \textbf{30} & $0.668306$ \\
-        \end{tabular}
-        \caption{user 32}
-    \end{subfigure}
-\end{table}
-
-The testing users' removed edges make it to the lists successfully; therefore, the removed friends are in the same community as the test users and can be recommended through friend suggestions.
-
-\subsection{Evaluation}
-
-The top-$N$ friend suggestions, generated for the testing users with Algorithm \ref{alg:friend_sug_com}, are evaluated against the three removed test edges.
-The ranked lists of friend-suggestions shown in Table \ref{tab:karateclub_emb_dists} are used to determine top-$N$ friend suggestions for each testing user and evaluate the recommendations by the hit-rate metric.
-The LOO-testing edges are printed bold in Table \ref{tab:karateclub_emb_dists}.
-The resulting top-$N$ hit-rates are plotted against the number of suggestions $N$ in Fig.
-\ref{fig:karateclub_hitrates}.
-
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width=.4\textwidth]{images/karateclub/hit_rates.png}
-    \caption{Hit-rates from top-$N$ recommendations on \textit{Zachary’s Karate Club graph}.}
-    \label{fig:karateclub_hitrates}
-\end{figure}
-
-Fig.
-\ref{fig:karateclub_hitrates} shows promising results for using ComE to generate top-$N$ friend suggestions on the small \textit{Zachary’s Karate Club graph}.
-The achieved hit-rate starts at 33\% for $N=1$, reaches 67\% at $N=3$, and maxes out at 100\% for $N=7$; therefore, when presenting three friend suggestions to one test user, at least in $66\%$ of cases a friend suggestion would be of interest.
-
-\section{Facebook Friend Suggestions}
-
-The recommender system based on community embeddings with the Bayesian Gaussian mixture model and variational inference is quantitatively evaluated.
-Friend suggestions are generated for the \textit{Social Circles: Facebook} dataset \cite{social_circles_fb, social_circles_in_ego_networks}.
-
-The method of generating friend suggestions using ComE BGMM+VI embeddings and the smallest-distance approach is applied to the \textit{Social Circles: Facebook} graph.
-Both Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} are evaluated.
-The top-$N$ recommendations are evaluated by the hit-rate metric.
-
-\subsection{Dataset}
-
-The \textit{Social Circles: Facebook} dataset was sourced from the profile and network data of ten ego-networks (egonets) through a custom Facebook app by \citeauthor{social_circles_in_ego_networks} as part of their \citeyear{social_circles_in_ego_networks} article \textit{Learning to Discover Social Circles in Ego Networks} \cite{social_circles_in_ego_networks}.
-
-Fig.
-\ref{fig:egonet} shows an illustration of the ego-network of user $u$.
-User $u$ has direct connections to $v_i$, called alters.
-Alters can be connected to each other and build circles of users, like families, high school friends, or college friends.
-
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width=0.48\textwidth]{images/fb/egonet.png}
-    \caption{Illustrative example of an ego-network \cite{social_circles_in_ego_networks}.}
-    \label{fig:egonet}
-\end{figure}
-
-\citeauthor{social_circles_in_ego_networks} obtained profile and network data from ten Facebook ego-networks. The dataset contains 4039 users (nodes) and 88234 friendships (edges).
-
-\subsection{Training and Testing Split}
-
-The \textit{Social Circles: Facebook} graph is split into training and testing parts.
-For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.
-Users with only one connection are removed from the dataset since no network information is added by them.
-
-$1169$ of the 4039 users of the \textit{Social Circles: Facebook} graph have at least $50$ connections and are used for testing friend suggestions.
-The threshold of $50$ friends for considering users for testing is set to make sure enough data is available to make sensable suggestions for testing users.
-
-Now that the testing users are identified, the leave-one-out method is employed for obtaining one testing friendship per testing user.
-The friendship chosen to leave out for testing for each testing user is chosen randomly using NumPy's random choice \cite{oliphant2006guide, van2011numpy}.
-
-\subsection{Embedding}
-
-Community embeddings for the \textit{Social Circles: Facebook} graph are generated by running ComE BGMM+VI with the hyperparameters listed in Table \ref{tab:params_fb}.
-
-\begin{table}
-    \centering
-    \caption{Hyperparameters used for the \textit{Social Circles: Facebook} dataset.}
-    \label{tab:params_fb}
-    \begin{tabular}{ l | c | r }
-        parameter                    & notation & value       \\
-        \hline
-        \hline
-        number\_walks                & $\gamma$ & $10$        \\
-        \hline
-        walk\_length                 & $\ell$   & $80$        \\
-        \hline
-        representation\_size         & $D$      & $2, 4, 128$ \\
-        \hline
-        num\_workers                 & \empty   & $10$        \\
-        \hline
-        num\_iter                    & \empty   & $3$         \\
-        \hline
-        com\_n\_init                 & \empty   & $10$        \\
-        \hline
-        reg\_covar                   & \empty   & $0.00001$   \\
-        \hline
-        batch\_size                  & \empty   & $50$        \\
-        \hline
-        window\_size                 & $\zeta$  & $10$        \\
-        \hline
-        negative                     & $m$      & $5$         \\
-        \hline
-        lr                           & \empty   & $0.025$     \\
-        \hline
-        alpha                        & $\alpha$ & $0.1$       \\
-        \hline
-        beta                         & $\beta$  & $0.1$       \\
-        \hline
-        down\_sampling               & \empty   & $0.0$       \\
-        \hline
-        communities                  & $K$      & $10$        \\
-        \hline
-        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$   \\
-    \end{tabular}
-\end{table}
-
-As is shown in Table \ref{tab:params_fb}, the embedding's representation size $D$ is varied along three values.
-A representation size $D \in [2,4,128]$ is chosen to determine the effect of the embedding's dimensionality on friend suggestions.
-It is hypothesized, that two dimensions are not sufficient to capture the structure of the \textit{Social Circles: Facebook} graph, due to its size, therefore higher dimensional embeddings are evaluated as well.
-
-Fig.
-\ref{fig:fb_emb_anim_d2}, Fig.
-\ref{fig:fb_emb_anim_d4}, and Fig.
-\ref{fig:fb_emb_anim_d128} show the node embeddings generated by ComE BGMM+VI after one, two, and three iterations.
-The nodes are colored by community membership from ComE's community embedding and detection.
-The $4$- and $128$-dimensional node and community embeddings cannot effectively be plotted in two-dimensional print.
-Instead, in Fig.
-\ref{fig:fb_emb_anim_d4} and Fig.
-\ref{fig:fb_emb_anim_d128} the two first dimensions of the node and community embeddings are plotted to gain a narrow perspective on the resulting node embeddings.
-
-\newcommand{\fbEmbAnimD}[1]{
-    \begin{figure*}[htbp]
-        \centering
-        \begin{subfigure}{.33\textwidth}
-            \centering
-            \includegraphics[width=1\linewidth]{images/fb/anim_d#1_i0.png}
-            \caption{\nth{1} Iteration}
-        \end{subfigure}%
-        %
-        \begin{subfigure}{.33\textwidth}
-            \centering
-            \includegraphics[width=1\linewidth]{images/fb/anim_d#1_i1.png}
-            \caption{\nth{2} Iteration}
-        \end{subfigure}%
-        %
-        \begin{subfigure}{.33\textwidth}
-            \centering
-            \includegraphics[width=1\linewidth]{images/fb/anim_d#1_i2.png}
-            \caption{\nth{3} Iteration}
-        \end{subfigure}%
-        %
-        \caption{Node embeddings segment with representation size $D=#1$.}
-        \label{fig:fb_emb_anim_d#1}
-    \end{figure*}
-}
-
-\fbEmbAnimD{2}
-\fbEmbAnimD{4}
-\fbEmbAnimD{128}
-
-Fig.
-\ref{fig:fb_graphs} shows the graph spring plotted with networkx and nodes colored by community membership after the last iteration of ComE \cite{networkx}.
-The spring plotted graph structure on Fig.
-\ref{fig:fb_graphs} suggests, that with embedding dimension $D=4$, more fitting communities are detected than for $D=2$ and $D=128$.
-
-\begin{figure*}[htbp]
-    \centering
-    \begin{subfigure}{.33\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/graph_d2.png}
-        \caption{$D=2$}
-    \end{subfigure}%
-    %
-    \begin{subfigure}{.33\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/graph_d4.png}
-        \caption{$D=4$}
-    \end{subfigure}%
-    %
-    \begin{subfigure}{.33\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/graph_d128.png}
-        \caption{$D=128$}
-    \end{subfigure}%
-    %
-    \caption{Spring plotted graphs with nodes colored by community membership after last ComE BGMM+VI iteration.}
-    \label{fig:fb_graphs}
-\end{figure*}
-
-Fig.
-\ref{fig:fb_emb_anim_d2} shows node and community embeddings with representation size $D=2$.
-The limiting two-dimensional space does not allow ComE to represent the \textit{Social Circles: Facebook} graph's structure sufficiently to utilize the $K=10$ components of the Gaussian mixture model.
-Only three components are used after the last of three iterations.
-
-While Fig.
-\ref{fig:fb_emb_anim_d4} and Fig.
-\ref{fig:fb_emb_anim_d128} show only the first two dimensions, of the node embedding of 4 or 128 dimensionality respectively, the general structure of communities is visible.
-The development of the node embeddings over the three iterations shows the manifestation of the first-, second-, and higher-order proximity by ComE's closed loop.
-The node embeddings are densified in their communities and spread out in regards to other communities.
-All $K=10$ components of the Gaussian mixture model are used for $D=4$ and $D=128$.
-
-The nodes are distributed over the $K=10$ communities as shown in the histograms in Fig.
-\ref{fig:fb_emb_weights}.
-Fig.
-\ref{fig:fb_emb_weights_d2} shows that with representation size $D=2$, ComE BGMM+VI drops 6 communities, effecitvelly using only 4 communities in the final embedding, because the two-dimensional node embeddings are not expressive enough to support more communities.
-Fig.
-\ref{fig:fb_emb_weights_d4} and Fig.
-\ref{fig:fb_emb_weights_d128} show that with representation sizes $D=4$ and $D=128$ respectivelly, all $K=10$ communities are used.
-
-The tradeoff between the representation size of the embedding $D$ and the embedding expressiveness is demonstrated.
-When choosing a too low-dimensional embedding, the embedding might not be expressive enough to capture the communities and reduces the quality of the community embedding and detection.
-When choosing a too high-dimensional embedding, the node embeddings get so expressive that BGMM+VI is not able to drop unused communities because all communities are used in one of the many dimensions.
-
-\begin{figure*}[htbp]
-    \centering
-    %
-    \begin{subfigure}{.33\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/weights_d2.png}
-        \caption{$D=2$}
-        \label{fig:fb_emb_weights_d2}
-    \end{subfigure}%
-    %
-    \begin{subfigure}{.33\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/weights_d4.png}
-        \caption{$D=4$}
-        \label{fig:fb_emb_weights_d4}
-    \end{subfigure}%
-    %
-    \begin{subfigure}{.33\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/weights_d128.png}
-        \caption{$D=128$}
-        \label{fig:fb_emb_weights_d128}
-    \end{subfigure}%
-    %
-    \caption{Weights of communities with $K=10$ and $D \in [2,4,128]$.}
-    \label{fig:fb_emb_weights}
-\end{figure*}
-
-\subsection{Friend Suggestions}
-
-With node and community embeddings and community assignments generated for $D \in [2,4,128]$ and $K=10$ with training data in the previous section, top-$N$ friend suggestion recommendations are generated from the training node and community embeddings with Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com}.
-
-The shortest-embeddings-distance approach is employed, recommending friends to users who have the shortest distance node embeddings over all users (Algorithm \ref{alg:friend_sug}) or only over users in the same community (Algorithm \ref{alg:friend_sug_com}).
-Users who are already connected are not recommended again.
-
-\section{Evaluation}
-
-The charts in Fig. \ref{fig:fb_hit_rates} show the hit-rates for top-$N$ recommendations for testing users.
-
-\begin{figure}[htbp]
-    \centering
-    % d2
-    \begin{subfigure}{.48\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/hit_rates_d-2.png}
-        \caption{$D=2$ on linear $N$-axis.}
-        \label{fig:fb_hit_rates_d2}
-    \end{subfigure}%
-
-    % d4
-    \begin{subfigure}{.48\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/hit_rates_d-4.png}
-        \caption{$D=4$ on linear $N$-axis.}
-        \label{fig:fb_hit_rates_d4}
-    \end{subfigure}%
-
-    % d128
-    \begin{subfigure}{.48\textwidth}
-        \centering
-        \includegraphics[width=1\linewidth]{images/fb/hit_rates_d-128.png}
-        \caption{$D=128$ on linear $N$-axis.}
-        \label{fig:fb_hit_rates_d128}
-    \end{subfigure}%
-
-    \caption{Hit-rates of \textit{Social Circles: Facebook} graph's top-$N$ friend suggestions on $N$-axis for $D \in [2,4,128]$.
-Hit-rate lines for Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} overlap for $D=2$ and $D=4$.}
-    \label{fig:fb_hit_rates}
-\end{figure}
-
-Two hit-rate lines are shown for each embedding representation size $D \in [2,4,128]$ on $N$-axis.
-The blue line shows the hit-rates when generating friend suggestions from all users (Algorithm \ref{alg:friend_sug}).
-The orange line shows the hit-rate when generating recommendations from users in the same community (Algorithm \ref{alg:friend_sug_com}).
-For lower-dimensional embeddings with $D \in [2,4]$, the hit-rates for Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} largely overlap.
-
-Table \ref{tab:fb_hit_rates} shows readings of hit-rates over the algorithms and embedding dimensionality for top-$5$, top-$10$, and top-$20$ recommendations:
-
-\begin{figure}[htbp]
-    \centering
-    \caption{Hit-rates of \textit{Social Circles: Facebook} graph's top-$N$ friend suggestions for algorithms \ref{alg:friend_sug} and \ref{alg:friend_sug_com}, $N \in [5, 10, 20]$, and $D \in [2, 4, 128]$.}
-    \label{tab:fb_hit_rates}
-    \begin{tabular}{ l | l | c | c | c }
-        \hline
-        \empty                                              & \empty  & top-$5$         & top-$10$        & top-$20$        \\
-        \hline
-        \hline
-        \multirow{3}{*}{Algorithm \ref{alg:friend_sug}}     & $D=2$   & 8.5\%           & 14.2\%          & 21.9\%          \\
-                                                            & $D=4$   & 18.1\%          & 26.5\%          & 38.4\%          \\
-                                                            & $D=128$ & \textbf{31.3\%} & \textbf{43.6\%} & \textbf{58.8\%} \\
-        \hline
-        \multirow{3}{*}{Algorithm \ref{alg:friend_sug_com}} & $D=2$   & 8.4\%           & 14.1\%          & 21.8\%          \\
-                                                            & $D=4$   & 18.0\%          & 26.4\%          & 38.2\%          \\
-                                                            & $D=128$ & \textbf{24.1\%} & \textbf{32.9\%} & \textbf{41.8\%} \\
-        \hline
-    \end{tabular}
-\end{figure}
-
-Hit-rates based on top-$5$, top-$10$, and top-$20$ recommendations are chosen since higher $N$ are not practical for recommendations.
-
-Presenting only the top-$10$ friend suggestions to users, a hit-rate of $43.6\%$ when considering all users, or $32.9\%$, when considering only users from the same community, is achieved.
-A top-$5$ hit-rate of $31.3\%$ (all users) and $24.1\%$ (same community) is achieved, which can be considered viable for business use-cases, especially for only showing 5 recommendations to users.
-A top-$20$ hit-rate of $58.8\%$ (all users) and $41.8\%$ (same community) adds to the presented results.
-
-\section{Conclusion}
-
-The proposed shortest-embedding-distance friend suggestion recommender system based on embeddings generated with the community embeddings algorithm ComE BGMM+VI shows promising results.
-A top-$10$ recommendations hit-rate of $43.6\%$ when considering all users (Algorithm \ref{alg:friend_sug}), and over $32.9\%$ when considering users in the same community (Algorithm \ref{alg:friend_sug_com}) is achieved for Facebook friend suggestions.
-Both results were achieved with a representation size $D=128$.
-Lower hit-rates are obtained for $D=2$ and $D=4$.
-See Table \ref{tab:fb_hit_rates}.
-
-Algorithm \ref{alg:friend_sug} achieves the better hit-rate in comparison to Algorithm \ref{alg:friend_sug_com}, since all users are considered for embedding instead of just users of the same community.
-Algorithm \ref{alg:friend_sug_com}, on the other hand, has better time complexity than Algorithm \ref{alg:friend_sug}, since only users in the same community are considered for friend suggestions instead of all users.
-Algorithm \ref{alg:friend_sug_com} is scalable to larger datasets if the number of communities $K$ can be considered to scale with the graph size $(\abs{V}, \abs{E})$.
-
-\subsection{Future Work}
-
-This work explores generating friend suggestions from node and community embeddings generated by ComE.
-Three values of the representation size $D$ were used in experimentation.
-Experimenting with more, different combinations of hyperparameters, and especially varying the number of communities $K$ might be of interest.
-Since in this paper, friend suggestions were explored, it is left for future work to explore item-based recommendations using embeddings.
-More community-centric recommendations, which deeply consider the community embeddings in generating recommendations, are left for future work.
-
 %%% Bibliography

### 2020-10-14 paper-ffs.tex
write intro

@@
 
+Graphs, such as social networks, knowledge graphs, content-rating graphs, and communication networks, emerge naturally from various real-world situations.
+Analyzing these graphs leads to findings and understanding of the underlying structures, coherences, and dependencies.
+Recently, methods for embedding graph's nodes into lower-dimensional Euclidean spaces, called graph embeddings, have gained traction in multiple areas of data science research.\cite{Goyal_2018}
+
+Community Embeddings, in addition to embedding a graph's nodes through first- and second-order proximity, also preserve higher-order proximity by embedding clusters present in the graph data.
+The graph and community embedding algorithm ComE aims to preserve first-, second- and higher-order proximity by embedding a graph's nodes and communities.\cite{ComE} ComE requires prior knowledge of the number of communities $K$.
+In this paper, ComE is extended to utilizing a Bayesian Gaussian mixture model with variational inference for learning community embeddings (ComE BGMM+VI), similar to ComE+ published by \citeauthor{ComE+} in \citeyear{ComE+}.\cite{ComE+} ComE BGMM+VI takes $K$ as the maximum number of communities and drops components through a trade-off hyperparameter.
+
+The recent \citeyear{ComE} graph embeddings algorithm ComE is extended similarily to the \citeyear{ComE+} ComE+ by taking a Bayesian approach.
+The open-source code for the Bayesian approach to ComE's community embedding is published on GitHub\footnote{at \citeurl{ComE_BGMM_GH}} and serves as a contribution to community embedding research.\cite{ComE_BGMM_GH}
+
+The original ComE paper \citetitle{ComE} by \citeauthor{ComE} and the ComE+ paper \citetitle{ComE+} by \citeauthor{ComE+} in combination with the ComE source code have provided the basis and architecture for the community embeddings utilized in this work.\cite{ComE, ComE+, ComE_GH}
+
+Multiple surveys and articles on graph embeddings were consulted to build a full picture of the current state of graph embedding research.
+Especially the \citeyear{Goyal_2018} survey \citetitle{Goyal_2018} by \citeauthor{Goyal_2018} and the \citeyear{rossi20tkdd-roles} paper \citetitle{rossi20tkdd-roles} by \citeauthor{rossi20tkdd-roles} have proven to be primary resources for understanding the current landscape of graph embedding research.\cite{Goyal_2018, rossi20tkdd-roles}
+
+On part of comparing the Bayesian Gaussian mixture model with variational inference to the Gaussian mixture model with expectation-maximization, the \citeyear{Bishop06} book \citetitle{Bishop06} by \citeauthor{Bishop06} includes essential statistics and information science knowledge and explanations.\cite{Bishop06}
+
 %%% Bibliography

### 2020-10-14 paper-ComE-BGMM-VI.tex
rename file
cite . format

--- 2020-10-14 paper-ffs.tex
@@
 
+Graphs, such as social networks, knowledge graphs, content-rating graphs, and communication networks, emerge naturally from various real-world situations.
+Analyzing these graphs leads to findings and understanding of the underlying structures, coherences, and dependencies.
+Recently, methods for embedding graph's nodes into lower-dimensional Euclidean spaces, called graph embeddings, have gained traction in multiple areas of data science research \cite{Goyal_2018}.
+
+Community Embeddings, in addition to embedding a graph's nodes through first- and second-order proximity, also preserve higher-order proximity by embedding clusters present in the graph data.
+The graph and community embedding algorithm ComE aims to preserve first-, second- and higher-order proximity by embedding a graph's nodes and communities\cite{ComE}.
+ComE requires prior knowledge of the number of communities $K$.
+In this paper, ComE is extended to utilizing a Bayesian Gaussian mixture model with variational inference for learning community embeddings (ComE BGMM+VI), similar to ComE+ published by \citeauthor{ComE+} in \citeyear{ComE+} \cite{ComE+}.
+ComE BGMM+VI takes $K$ as the maximum number of communities and drops components through a trade-off hyperparameter.
+
+The recent \citeyear{ComE} graph embeddings algorithm ComE is extended similarily to the \citeyear{ComE+} ComE+ by taking a Bayesian approach.
+The open-source code for the Bayesian approach to ComE's community embedding is published on GitHub\footnote{at \url{https://github.com/abegehr/ComE_BGMM}} and serves as a contribution to community embedding research \cite{ComE_BGMM_GH}.
+
+The original ComE paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs} by \citeauthor{ComE} and the ComE+ paper \textit{Embedding Both Finite and Infinite Communities on Graphs} by \citeauthor{ComE+} in combination with the ComE source code have provided the basis and architecture for the community embeddings utilized in this work \cite{ComE, ComE+, ComE_GH}.
+
+Multiple surveys and articles on graph embeddings were consulted to build a full picture of the current state of graph embedding research.
+Especially the \citeyear{Goyal_2018} survey \textit{Graph Embedding Techniques, Applications, and Performance: A Survey} by \citeauthor{Goyal_2018} and the \citeyear{rossi20tkdd-roles} paper \textit{On Proximity and Structural Role-based Embeddings in Networks: Misconceptions, Techniques, and Applications} by \citeauthor{rossi20tkdd-roles} have proven to be primary resources for understanding the current landscape of graph embedding research \cite{Goyal_2018, rossi20tkdd-roles}.
+
+On part of comparing the Bayesian Gaussian mixture model with variational inference to the Gaussian mixture model with expectation-maximization, the \citeyear{Bishop06} book \textit{Pattern Recognition and Machine Learning (Information Science and Statistics)} by \citeauthor{Bishop06} includes essential statistics and information science knowledge and explanations \cite{Bishop06}.
+
+\section{Community Embedding}
+
 %%% Bibliography

### 2020-10-14 paper-ComE-BGMM-VI.tex
do sec1 ComE

@@
 \usepackage{url}
+\usepackage{tabularx}
 
@@
 
+\section{Graph Embedding}
+
+A graph embedding is a representation of a graph data structure in lower-dimensional space.
+Utilizing graph embeddings has recently gained traction in the research community for representing and analyzing graph data \cite{Goyal_2018}.
+Possible applications of graph embeddings include node and graph classification, anomaly detection, link prediction, recommender systems, graph compression, and visualizations \cite{rossi20tkdd-roles}.
+
+There are multiple advantages of graph embeddings over the original graph data-structure $G = (V, E)$:
+
+\begin{enumerate}
+    \item \textbf{Machine learning algorithms on graphs are limited \cite{Godec2018}}.Only specific mathematics, statistics, and machine learning algorithms can be applied to the specific graph data-structure $G$ consisting of nodes $V$ and edges $E$.
+    \item \textbf{Vector operations are simpler and faster than comparable graph operations \cite{Godec2018}.} A graph embedding is a representation of a graph in lower-dimensional space of dimension $d$, therefore nodes are assigned a feature vector of size $d$, which can then be operated on using vector operations.
+    \item \textbf{Embeddings are compressed representations \cite{Godec2018}.} A trivial feature vector for a node $v \in V$ of the graph $G = (V, E)$ is of length $\abs{V}$ with an entry for each node $v_i \in V$ valued by the existing or non-existing, binary or weighted edge.
+When determining this trivial representation for all nodes $v \in V$, the adjacency matrix of size $\abs{V} \times \abs{V}$ is obtained.
+An embedding of the same graph $G = (V, E)$ is obtained through embedding $G$ into a $d$-dimensional space.
+The resulting embedding is of size $\abs{V} \times d$.
+An embedding is performed into a lower-dimensional space, therefore $d \ll \abs{V}$.
+\end{enumerate}
+
+Popular graph embedding algorithms include DeepWalk, introduced by \citeauthor{Perozzi2014DeepWalkOL} in their \citeyear{Perozzi2014DeepWalkOL} paper \textit{DeepWalk}, and Node2Vec, introduced by \citeauthor{Grover2016node2vecSF} in their \citeyear{Grover2016node2vecSF} paper \textit{node2vec: Scalable Feature Learning for Networks} \cite{Perozzi2014DeepWalkOL, Grover2016node2vecSF}.
+Both DeepWalk and Node2Vec demonstrate competitive results for the use-case of multi-label classification and are promising options for graph embeddings.
+
 \section{Community Embedding}
+
+Recently, a graph embedding algorithm was altered and extended to include the concept of communities.
+Communities refer to groups of nodes that have high inner connectivity and low outer connectivity to other communities \cite{ComE}.
+The proximity of nodes hereby is extended to include the concept of higher-order proximity, in addition to first-order and second-order-proximity \cite{ComE}.
+
+\citeauthor{ComE} developed the Community Embedding algorithm ComE in their \citeyear{ComE} paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs} \cite{ComE}.
+
+\subsection{ComE Algorithm}
+\label{sec:ComE_alg}
+
+ComE consists of three parts: node embedding, community detection, and community embedding.
+A closed-loop relationship between node embedding, community detection, and community embedding is indicated by \citeauthor{ComE}.
+The closed-loop is exploited in an iterative algorithm optimizing the node and community embeddings.
+
+The ComE algorithm is sketched out as follows:
+
+\begin{enumerate}
+    \item Sample the graph by executing random walks.
+    \item Initialize the node embedding $\Phi$ and context embedding $\Phi'$ with the node embedding algorithm DeepWalk\cite{Perozzi2014DeepWalkOL} with random walks.
+    \item Fit a Gaussian mixture model (GMM) representing communities to $(\Phi, \Phi')$ using the expectation-maximization (EM) algorithm, while keeping $(\Phi, \Phi')$ fixed.
+Yield the parameters of the GMM as the community embedding: mixed community membership $\Pi$, Gaussian means $\Psi$, and Gaussian covariances $\Sigma$.
+    \item Use stochastic decent with first-, second-, and higher-order proximity optimization functions to adjust $(\Phi, \Phi')$ to the Gaussian mixture model, keeping the GMM's parameters $(\Pi, \Psi, \Sigma)$ fixed.
+Yield the next node embedding and context embedding $(\Phi, \Phi')$ updated by first-, second-, and higher-order proximity.
+    \item Repeat steps 3 and 4 at will. Step 3 represents community detection and embedding. Step 4 represents updating the node embedding.
+\end{enumerate}
+
+The formulas and concepts used in this chapter are borrowed from the original ComE paper.\cite{ComE} For a pseudocode implementation including more detail, see algorithm 1 of the original ComE paper \cite{ComE}.
+\citeauthor{ComE_GH} published an implementation of the ComE algorithm on GitHub\footnote{at \url{https://github.com/andompesta/ComE}}.
+
+\subsection{ComE Hyperparameters}
+\label{sec:ComE_params}
+
+ComE requires multiple hyperparameters to be set prior to computation. The following table lists the hyperparameters, a description for each, and the notation used in literature:
+
+\begin{table}[htbp]
+    \centering
+    \caption{Hyperparameters used by \citeauthor{ComE_GH}'s implementation of ComE \cite{ComE_GH}.}
+    \label{table:params_ComE}
+    \begin{tabularx}{\linewidth}{ l | c | X }
+        parameter            & notation & description                                                                     \\
+        \hline
+        \hline
+        number\_walks        & $\gamma$ & number of random walks for each node                                            \\
+        \hline
+        walk\_length         & $\ell$   & length of each walk                                                             \\
+        \hline
+        representation\_size & $D$      & dimensionality of the embedding                                                 \\
+        \hline
+        num\_workers         & \empty   & number of threads                                                               \\
+        \hline
+        num\_iter            & \empty   & number of overall iterations                                                    \\
+        \hline
+        com\_n\_init         & \empty   & number of initializations to run on the communtiy embedding model (GMM or BGMM) \\
+        \hline
+        reg\_covar           & \empty   & regularization coefficient for ensuring positive covariance                     \\
+        \hline
+        batch\_size          & \empty   & size of the batch                                                               \\
+        \hline
+        window\_size         & $\zeta$  & windows size used to compute the context embedding                              \\
+        \hline
+        negative             & $m$      & number of negative samples                                                      \\
+        \hline
+        lr                   & \empty   & learning rate                                                                   \\
+        \hline
+        alpha                & $\alpha$ & Trade-off parameter for context embedding                                       \\
+        \hline
+        beta                 & $\beta$  & Trade-off parameter for community embedding                                     \\
+        \hline
+        down\_sampling       & \empty   & perform down sampling of common nodes                                           \\
+        \hline
+        communities          & $K$      & number of communities                                                           \\
+    \end{tabularx}
+\end{table}
+
+\subsection{Non-Bayesian and Bayesian Gaussian mixture models}
+
+\begin{figure}[htbp]
+    \centering
+    \begin{subfigure}[b]{.4\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/ComE/GMM.png}
+        \caption{Non-Bayesian}
+    \end{subfigure}%
+    \begin{subfigure}[b]{.5\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/ComE/BGMM.png}
+        \caption{Bayesian}
+    \end{subfigure}%
+    \caption{Gaussian mixture models in plate notation\protect\footnotemark.}
+    \label{fig:GMM_vs_BGMM_plates}
+\end{figure}
+\footnotetext{By Benwing - Created using LaTeX, TikZ, CC BY 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=18934624}}
+
+Figure \ref{fig:GMM_vs_BGMM_plates} compares a non-Bayesian and a Bayesian GMM side-by-side in plate notation: Squares indicate fixed parameters and circles indicate random variables.
+
+The Bayesian Gaussian mixture model assumes all parameters of the Gaussian mixture model (number of communities $K$, means $\mu_k$, and covariances $\sigma_k$) are themselves generated by some distribution.
+This is represented in the plate diagrams in Figure \ref{fig:GMM_vs_BGMM_plates}.
+
+A Bayesian Gaussian mixture model (BGMM) with variational inference (VI) offers a solution to the issue of introducing the number of communities $K$ as a fixed parameter, by taking advantage of Bayesian methods.
+
+A BGMM with VI tends to utilize some components heavily while eliminating other components, if the provided dataset suggests so, by applying a tradeoff parameter.
+Therefore, the parameter $K$ represents a maximum of communities and not the final number of communities like for GMMs.
+
+\subsection{Algorithm ComE BGMM+VI}
+\label{sec:ComE_BGMM_alg}
+
+The algorithm can stay mostly unchanged to the algorithm sketched out in Section \ref{sec:ComE_alg}.
+Only steps 3 and 4 need to be adjusted to handle community detection and embedding and node embeddings using a Bayesian Gaussian mixture model (BGMM) with variational inference (VI) instead of a non-Bayesian Gaussian mixture model (GMM) with expectation-maximization (EM).
+
+A Bayesian Gaussian mixture model, just like a non-Bayesian Gaussian mixture model, produces mixed community membership $\Pi$, which means $\Psi$, and covariances $\Sigma$ for communities based on node embeddings $\Phi$.
+Therefore, the inputs and outputs of step 3 stay the same.
+
+In their \citeyear{ComE+} paper \textit{Embedding Both Finite and Infinite Communities on Graphs}, \citeauthor{ComE+} present their version of ComE with an infinite Bayesian-approach to community detection and embedding.
+As of the time of writing of this paper, no source code for ComE+ was published.
+While ComE+ and ComE BGMM+VI both take a Bayesian approach to community embedding, ComE+ redefined the optimization function for higher-order proximity based on the Bayesian Gaussian mixture model, while ComE BGMM+VI simply uses a BGMM for community embedding and detection and leaves the node embedding step unchanged.
+
+The implementation of ComE BGMM+VI\footnote{ComE BGMM+VI is published at \url{https://github.com/abegehr/ComE_BGMM} \cite{ComE_BGMM_GH}.} presented and used in this paper, utilizes sklearn's implementation of the Bayesian Gaussian mixture model and variational inference references at \textit{sklearn.mixture.BayesianGaussianMixture}.\cite{scikit-learn, sklearn_api}
+
+\subsection{Hyperparameters BGMM}
+\label{ComE_BGMM_params}
+
+Switching from the original ComE's GMM to a Bayesian GMM, will require one additional parameter: the BGMM's weight concentration prior, or simply concentration.
+
+The BGMM's concentration is a trade-off parameter that directly influences the number of components utilized for community embedding.
+A higher concentration parameter leads to more components active.
+A lower concentration parameter leads to fewer components active.
+
+ComE with BGMM and VI has the same hyperparameters as ComE with GMM and EM, as presented in Section \ref{sec:ComE_params}, and in addition also the following:
+
+\begin{table}[htbp]
+    \centering
+    \caption{Hyperparameter added to ComE for supporting a BGMM for community embedding \cite{ComE_BGMM_GH}.}
+    \label{table:params_ComE_BGMM}
+    \begin{tabularx}{\linewidth}{ l | c | X }
+        parameter                    & notation & description                               \\
+        \hline
+        \hline
+        weight\_concentration\_prior & $\Gamma$ & Bayesian GMM's weight concentration prior \\
+    \end{tabularx}
+\end{table}
 

### 2020-10-14 paper-ComE-BGMM-VI.tex
do visual example

@@
 
+\section{Visual Example}
+\label{ch:come_visual}
+
+In this section community embeddings based on a standard Gaussian mixture model with expectation-maximization (GMM+EM) and community embeddings based on a Bayesian Gaussian mixture model with variational inference (BGMM+VI) are compared visually.
+Two-dimensional visualizations are computed on a small graph-dataset using multiple values for the hyperparameter number of communities $K$ to show the advantage of BGMM+VI over GMM+EM under specific conditions.
+
+\subsection{Karate Club Dataset}
+\label{sec:karate_club}
+
+\textit{Zachary’s Karate Club graph} is a popular graph dataset for visualizing graph algorithms, due to its small size \cite{KarateClubDS}.
+
+The data was collected from a university karate club by \citeauthor{Zachary1977AnIF} as part of the \citeyear{Zachary1977AnIF} paper \textit{Zachary’s Karate Club graph}.
+The small Karate Club graph consists of 34 nodes, each representing one person as part of the karate club.
+The presence of an edge represents a social tie among two members of the group.
+The absence of an edge represents no social tie among the two members of the group.
+Edges are bidirectional.
+\textit{Zachary’s Karate Club graph} provided a binary version (ZACHE) and a weighted version (ZACHC) of the graph.
+We will consider the binary version of the karate club graph.
+
+The following hyperparameters are used for computing both ComE with GMM and BGMM results:
+
+\begin{table}[H]
+    \centering
+    \caption{Hyperparameters used for the Karate Club dataset. See Table \ref{table:params_ComE} for descriptions of hyperparameters.}
+    \label{table:params_KCvisual}
+    \begin{tabular}{ l | c | r }
+        parameter                    & notation & value        \\
+        \hline
+        \hline
+        number\_walks                & $\gamma$ & $10$         \\
+        \hline
+        walk\_length                 & $\ell$   & $80$         \\
+        \hline
+        representation\_size         & $D$      & $2$          \\
+        \hline
+        num\_workers                 & \empty   & $10$         \\
+        \hline
+        num\_iter                    & \empty   & $1$          \\
+        \hline
+        reg\_covar                   & \empty   & $0.00001$    \\
+        \hline
+        batch\_size                  & \empty   & $50$         \\
+        \hline
+        window\_size                 & $\zeta$  & $10$         \\
+        \hline
+        negative                     & $m$      & $5$          \\
+        \hline
+        lr                           & \empty   & $0.025$      \\
+        \hline
+        alpha                        & $\alpha$ & $0.1$        \\
+        \hline
+        beta                         & $\beta$  & $0.1$        \\
+        \hline
+        down\_sampling               & \empty   & $0.0$        \\
+        \hline
+        communities                  & $K$      & $[2,3,4,15]$ \\
+        \hline
+        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$    \\
+    \end{tabular}
+\end{table}
+
+For each $K \in [2,3,4,15]$ embeddings based on ComE with GMM and ComE with BGMM are computed.
+Each combination of a $K$ and choice of a mixture model for community detection and embedding, three charts are obtained:
+
+\begin{enumerate}
+    \item Graph: nodes positioned using the Fruchterman-Reingold force-directed (spring) algorithm and colored by community assignment.
+    \item Embeddings: obtained node and community embeddings.
+Ellipses represent communities and the underlying Gaussian component.
+Points colored by the community assignment represent node embeddings.
+    \item Weights: bar chart of weight for each community.
+\end{enumerate}
+
+\newcommand{\visresults}[2]{
+    \begin{figure}[H]
+        \centering
+        \begin{subfigure}{.15\textwidth}
+            \centering
+            \includegraphics[width=1\linewidth]{images/results/#1/graph_#2.png}
+            \caption{Graph}
+        \end{subfigure}%
+        %
+        \begin{subfigure}{.15\textwidth}
+            \centering
+            \includegraphics[width=1\linewidth]{images/results/#1/embeddings_#2.png}
+            \caption{Embeddings}
+        \end{subfigure}%
+        %
+        \begin{subfigure}{.15\textwidth}
+            \centering
+            \includegraphics[width=1\linewidth]{images/results/#1/weights_#2.png}
+            \caption{Weights}
+        \end{subfigure}%
+        %
+        \caption{ComE with #1, $K=#2$}
+        \label{fig:results#1#2}
+    \end{figure}
+}
+
+\subsection{K=2}
+
+\visresults{GMM}{2}
+\visresults{BGMM}{2}
+
+Figure \ref{fig:resultsGMM2} and Figure \ref{fig:resultsBGMM2}.
+Two is the intended number of communities.
+Setting $K=2$ leads to similar results for both using a GMM and a BGMM for community detection and embedding.
+Although the labels are inverted, and the separation between both community assignments is slightly different, the general community assignment and structure are the same.
+
+\subsection{K=3}
+
+\visresults{GMM}{3}
+\visresults{BGMM}{3}
+
+Figure \ref{fig:resultsGMM3} and Figure \ref{fig:resultsBGMM3}.
+Choosing $K=3$ leads to worse results when using a GMM in comparison to the BGMM.
+The GMM utilizes the three components, while the BGMM drops the unnecessary third component and utilizes only the two necessary components.
+
+\subsection{K=4}
+
+\visresults{GMM}{4}
+\visresults{BGMM}{4}
+
+Figure \ref{fig:resultsGMM4} and Figure \ref{fig:resultsBGMM4}.
+With $K=4$, the quality of embeddings based on the GMM further declines.
+The GMM utilizes all four communities.
+The result obtained by the BGMM also declines for this example, but wins over the GMM.
+The BGMM utilizes three communities.
+
+\subsection{K=15}
+
+\visresults{GMM}{15}
+\visresults{BGMM}{15}
+
+Figure \ref{fig:resultsGMM15} and Figure \ref{fig:resultsBGMM15}.
+Providing $K=15$ at initialization further underlines the potential advantage of BGMM over GMM.
+The GMM utilizes all 15 communities.
+It is clear that in this dataset of 34 nodes, no 15 communities are necessary for proper representation.
+The BGMM achieves a visually comparable quality to results at $K=2$ and $K=3$ by only utilizing 2 of the maximum 15 communities for representation.
+
+\subsection{Summary}
+
+As hypothesized, the community embedding algorithm ComE underperforms when the hyperparameter number of communities $K$ is unknown.
+A Bayesian approach with variational inference similar to ComE+ is proposed to use the tradeoff parameter weight concentration before dropping unused communities if $K$ is chosen too high.
+
+The small Karate Club dataset, consisting only of 34 nodes, is used to test the hypothesis through a visual comparison experiment.
+The Karate Club graph is embedded into two-dimensional Euclidean space with ComE and ComE BGMM+VI for multiple values for the hyperparameter number of communities $K = [2, 3, 4, 15]$.
+The hypothesized advantage of BGMM over GMM for unknown numbers of communities $K$ when embedding with ComE becomes clear.
+Even with $K=15$, ComE BGMM+VI identifies the $2$ communities, while the non-Bayesian ComE utilizes all $15$ communities.
+
 %%% Bibliography

### 2020-10-14 paper-ComE-BGMM-VI.tex
do ch quant eval

@@
 \usepackage{tabularx}
+\usepackage{longtable,multicol,makecell}
+\usepackage{pgfplots}
 
@@
     \centering
-    \caption{Hyperparameters used for the Karate Club dataset. See Table \ref{table:params_ComE} for descriptions of hyperparameters.}
+    \caption{Hyperparameters used for the Karate Club dataset.}
     \label{table:params_KCvisual}
@@
 \end{table}
+
+See Table \ref{table:params_ComE} for a descriptions of the hyperparameters.
 
@@
 
+\section{Quantitative Evaluation}
+\label{ch:come_eval}
+
+In their original \citeyear{ComE} paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs}, \citeauthor{ComE} identify three evaluation tasks for experimentation: graph visualization, community detection and node classification \cite{ComE}.
+Community detection and node classification can objectivelly be quantitatively evaluated, while graph visualization is subjectively evaluated.
+
+In this section, community embeddings based on a non-Bayesian Gaussian Mixture Model with Expectation-maximization (GMM+EM), as presented with the ComE algorithm, and community embeddings based on a Bayesian Gaussian Mixture Model with Variational Inference (BGMM+VI), similar to the ComE+ algorithm, are evaluated quantitatively over multiple values for the number of communities $K$ hyperparameter.
+
+\section{DBLP Dataset}
+
+To evaluate ComE GMM+EM and ComE BGMM+VI on the tasks graph visualization, community detection, and node classification, the DBLP dataset is considered:
+
+\begin{table}[H]
+    \centering
+    \caption{Dataset used for quantitative evaluation.}
+    \label{table:evaluation_datasets}
+    \begin{tabular}{ c || c | c | c | c }
+        \thead{graph                                                                            \\$G=(V, E)$} & \thead{\#(node)\\$\abs{V}$} & \thead{\#(edge)\\$\abs{E}$} & \thead{\#(class)\\$K$} & \thead{\#(label)/node} \\
+        \hline
+        \hline
+        \makecell{DBLP} & \makecell{13,184} & \makecell{48,018} & \makecell{5} & \makecell{$1$} \\
+    \end{tabular}
+\end{table}
+
+\subsection{Hyperparameters}
+\label{sec:come_eval_hyper-params}
+
+For comparability, the same hyperparameters are used for ComE GMM+EM and for the proposed ComE BGMM+VI, with the addition of the weight concentration prior hyperparameter for ComE BGMM+VI.
+The following hyperparameters were used for evaluating the DBLP dataset:
+
+\begin{table}[H]
+    \centering
+    \caption{Hyperparameters used for the DBLP dataset. See Table \ref{table:params_ComE} for descriptions of hyperparameters.}
+    \label{table:params_DBLP}
+    \begin{tabular}{ l | c | r }
+        parameter                    & notation & value         \\
+        \hline
+        \hline
+        number\_walks                & $\gamma$ & $10$          \\
+        \hline
+        walk\_length                 & $\ell$   & $80$          \\
+        \hline
+        representation\_size         & $D$      & $[128, 2]$    \\
+        \hline
+        num\_workers                 & \empty   & $10$          \\
+        \hline
+        num\_iter                    & \empty   & $3$           \\
+        \hline
+        com\_n\_init                 & \empty   & $10$          \\
+        \hline
+        reg\_covar                   & \empty   & $0.00001$     \\
+        \hline
+        batch\_size                  & \empty   & $50$          \\
+        \hline
+        window\_size                 & $\zeta$  & $10$          \\
+        \hline
+        negative                     & $m$      & $5$           \\
+        \hline
+        lr                           & \empty   & $0.025$       \\
+        \hline
+        alpha                        & $\alpha$ & $0.1$         \\
+        \hline
+        beta                         & $\beta$  & $0.1$         \\
+        \hline
+        down\_sampling               & \empty   & $0.0$         \\
+        \hline
+        communities                  & $K$      & $[2,5,10,20]$ \\
+        \hline
+        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$     \\
+    \end{tabular}
+\end{table}
+
+See Table \ref{table:params_ComE} for a descriptions of the hyperparameters.
+
+In addition to varying the hyper-parameter number of communities $K$, the dimensionality of the representation (representation\_size) is varied.
+The tasks are evaluated for two-dimensional and 128-dimensional embeddings to determine the effect of embedding size on the quality of results.
+
+\subsection{Community Detection}
+
+Community detection references the task of clustering nodes into communities and comparing the detected communities to the true communities as obtained from the true node classifications.
+The evaluation metrics are conductance and NMI.
+
+\subsubsection{Conductance}
+
+Figure \ref{fig:dblp_conductance} shows the conductance obtained from communities generated with ComE GMM+EM and ComE BGMM+VI on the DBLP graph with two- and 128-dimensional embeddings.
+Lower conductance is better.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{.5\linewidth}
+        \centering
+        \begin{tikzpicture}
+            \pgfplotstableread[row sep=\\,col sep=&]{
+                K	& GMM					& BGMM					\\
+                2	& 0.02054029917392275	& 0.021070234113712373	\\
+                5	& 0.030399077942659558	& 0.02551070276610302	\\
+                10	& 0.01722986509950581	& 0.030197444831591175					\\
+                20	& 0.029249335242380856	& 0.03289726818006926	\\
+            }\mydata
+            \begin{axis}[
+                    ybar,
+                    bar width=.5cm,
+                    width=\linewidth,
+                    height=\linewidth,
+                    legend style={at={(0.5,1)},
+                            anchor=north,legend columns=-1},
+                    symbolic x coords={2,5,10,20},
+                    xtick=data,
+                    point meta={y*100},
+                    yticklabel= {\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\%},
+                    ymin=0, ymax=0.15,
+                    xlabel={$K$},
+                    ylabel={Conductance},
+                    nodes near coords={\pgfmathprintnumber\pgfplotspointmeta\%},
+                    nodes near coords align={vertical},
+                ]
+                \addplot table[x=K,y=GMM]{\mydata};
+                \addplot table[x=K,y=BGMM]{\mydata};
+                \legend{GMM, BGMM}
+            \end{axis}
+        \end{tikzpicture}
+        \caption{two-dimensional embeddings.}
+        \label{fig:dblp_conductance_d2}
+    \end{subfigure}%
+    %
+    \begin{subfigure}{.5\linewidth}
+        \centering
+        \begin{tikzpicture}
+            \pgfplotstableread[row sep=\\,col sep=&]{
+                K	& GMM					& BGMM					\\
+                2	& 0.02072897576868255	& 0.05153027478804642	\\
+                5	& 0.013701852396354014	& 0.011492883939736796	\\
+                10	& 0.06247833021288399	& 0.04919872794163497					\\
+                20	& 0.11969825263644061	& 0.07018888966334426 \\
+            }\mydata
+            \begin{axis}[
+                    ybar,
+                    bar width=.5cm,
+                    width=\linewidth,
+                    height=\linewidth,
+                    legend style={at={(0.5,1)},
+                            anchor=north,legend columns=-1},
+                    symbolic x coords={2,5,10,20},
+                    xtick=data,
+                    point meta={y*100},
+                    yticklabel= {\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\%},
+                    ymin=0, ymax=0.15,
+                    xlabel={$K$},
+                    nodes near coords={\pgfmathprintnumber\pgfplotspointmeta\%},
+                    nodes near coords align={vertical},
+                ]
+                \addplot table[x=K,y=GMM]{\mydata};
+                \addplot table[x=K,y=BGMM]{\mydata};
+                \legend{GMM, BGMM}
+            \end{axis}
+        \end{tikzpicture}
+        \caption{128-dimensional embeddings.}
+        \label{fig:dblp_conductance_d128}
+    \end{subfigure}
+
+    \caption{Conductance on DBLP dataset. Lower conductance is better.}
+    \label{fig:dblp_conductance}
+\end{figure}
+
+As was expected, the conductance achieved with the true hyperparameter number of communities $K=5$ is very similar for GMM and BGMM.
+The conductance achieved for $K=2$ is similar for two-dimensional embeddings and higher for BGMM for 128-dimensional embeddings.
+
+The conductances for $K=10$ and $K=20$ are different depending on the dimensionality of the embedding.
+For a large representation size of 128 dimensions (see Figure \ref{fig:dblp_conductance_d128}), the conductance for the tested values for $K$ where $K>5$ increases in variability and decreases in overall quality in comparison to the smaller representation size of 2 dimensions (see Figure \ref{fig:dblp_conductance_d2}), where the conductance stays mostly the same with $K=10$ and $K=20$.
+
+Although ComE BGMM+VI produces better conductance than ComE GMM+EM for $K>5$, a clear advantage of BGMM over GMM for community embeddings can not be identified from this experiment alone.
+
+\subsubsection{Normalized Mutual Information}
+\label{sec:eval_nmi}
+
+Figure \ref{fig:dblp_nmi} shows the NMIs obtained from communities generated by ComE GMM+EM and ComE BGMM+VI with two- and 128-dimensional embeddings on the DBLP graph. Higher NMI is better.
+
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \begin{tikzpicture}
+            \pgfplotstableread[row sep=\\,col sep=&]{
+                K	& GMM					& BGMM					\\
+                2	& 0.4636647057169119	& 0.4623820822063423	\\
+                5	& 0.46257130393293455		& 0.46502447614120096		\\
+                10	& 0.43349289423082416		& 0.46005304584573076		\\
+                20	& 0.3999601374635516	& 0.45371067853180896	\\
+            }\mydata
+            \begin{axis}[
+                    ybar,
+                    bar width=.5cm,
+                    width=\linewidth,
+                    height=\linewidth,
+                    legend style={at={(0.5,1)},
+                            anchor=north,legend columns=-1},
+                    symbolic x coords={2,5,10,20},
+                    xtick=data,
+                    nodes near coords,
+                    nodes near coords align={vertical},
+                    ymin=0,ymax=1,
+                    xlabel={$K$},
+                    ylabel={NMI},
+                ]
+                \addplot table[x=K,y=GMM]{\mydata};
+                \addplot table[x=K,y=BGMM]{\mydata};
+                \legend{GMM, BGMM}
+            \end{axis}
+        \end{tikzpicture}
+        \caption{two-dimensional embeddings.}
+        \label{fig:dblp_nmi_d2}
+    \end{subfigure}%
+    %
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \begin{tikzpicture}
+            \pgfplotstableread[row sep=\\,col sep=&]{
+                K	& GMM					& BGMM					\\
+                2	& 0.4802479401582417	& 0.43575239496704926	\\
+                5	& 0.6746848015894669		& 0.668553614850714		\\
+                10	& 0.4963225850192131		& 0.47369605057376785		\\
+                20	& 0.3018911505699723	& 0.30816578800908423	\\
+            }\mydata
+            \begin{axis}[
+                    ybar,
+                    bar width=.5cm,
+                    width=\linewidth,
+                    height=\linewidth,
+                    legend style={at={(0.5,1)},
+                            anchor=north,legend columns=-1},
+                    symbolic x coords={2,5,10,20},
+                    xtick=data,
+                    nodes near coords,
+                    nodes near coords align={vertical},
+                    ymin=0,ymax=1,
+                    xlabel={$K$},
+                ]
+                \addplot table[x=K,y=GMM]{\mydata};
+                \addplot table[x=K,y=BGMM]{\mydata};
+                \legend{GMM, BGMM}
+            \end{axis}
+        \end{tikzpicture}
+        \caption{128-dimensional embeddings.}
+        \label{fig:dblp_nmi_d128}
+    \end{subfigure}
+
+    \caption{NMI on DBLP dataset. Higher NMI is better.}
+    \label{fig:dblp_nmi}
+\end{figure}
+
+ComE GMM+EM and ComE BGMM+VI perform very similar for the true hyperparameter number of communities $K=5$ over both two-dimensional and 128-dimensional embeddings, as was to be expected.
+The controlling experiment $K=2$ produces similar results for GMM and BGMM as well for both two- and 128-dimensional embeddings.
+
+ComE GMM+EM achieves an NMI of $0.675$ and ComE BGMM+VI achieves an NMI of $0.669$ for the 128-dimensional embedding.
+\citeauthor{ComE} state an NMI of over $0.7$ for the ComE algorithm \cite{ComE}.
+It is not clearly stated how many iterations are run specifically on DBLP in the original ComE paper, but it is stated that ComE converges within two or three iterations.\cite{ComE} In this experiment three iterations were applied, as stated in Section \ref{sec:come_eval_hyper-params}.
+
+With two-dimensional embeddings, an around $0.2$ units smaller NMI is achieved for $K=K^*=5$ than for the 128-dimensional embedding.
+It seems that the two-dimensional embedding is not capable to capture the complexity of the graph as well as the 128-dimensional embedding.
+
+The advantage of a smaller embedding becomes clearer for larger values of $K>K^*$.
+For $K=20$ the NMI for the 128-dimensional embedding is around $0.3$, while the BGMM two-dimensional embedding manages to hold at $0.45$.
+
+It seems that for higher $K>5$, ComE BGMM+VI can maintain higher NMI values for the two-dimensional embedding, while for the 128-dimensional embeddings, ComE BGMM+VI is not able to maintain the NMI score.
+Section \ref{sec:eval_DBLP_result_k} explores the different behavior of ComE BGMM+VI on low- and higher-dimensional embeddings further in comparison to ComE GMM+EM.
+
+\subsection{Number of Communities and Dimensionality}
+\label{sec:eval_DBLP_result_k}
+
+To understand why ComE BGMM+VI is able to maintain an NMI score for larger values of $K$ for the two-dimensional embedding, but not for the 128-dimensional embedding, the weights of the components for $K=20$ is presented:
+
+\begin{figure}[H]
+    \centering
+    % K=5
+    \begin{subfigure}[b]{.5\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/eval/weights_BGMM_d2_k5.png}
+        \caption{D=2, K=5}
+        \label{fig:dblp_weights_d2_k5}
+    \end{subfigure}%
+    %
+    \begin{subfigure}[b]{.5\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/eval/weights_BGMM_d128_k5.png}
+        \caption{D=128, K=5}
+        \label{fig:dblp_weights_d128_k5}
+    \end{subfigure}
+    % K=10
+    \begin{subfigure}[b]{.5\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/eval/weights_BGMM_d2_k10.png}
+        \caption{D=2, K=10}
+        \label{fig:dblp_weights_d2_k10}
+    \end{subfigure}%
+    %
+    \begin{subfigure}[b]{.5\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/eval/weights_BGMM_d128_k10.png}
+        \caption{D=128, K=10}
+        \label{fig:dblp_weights_d128_k10}
+    \end{subfigure}
+    % K=20
+    \begin{subfigure}[b]{.5\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/eval/weights_BGMM_d2_k20.png}
+        \caption{D=2, K=20}
+        \label{fig:dblp_weights_d2_k20}
+    \end{subfigure}%
+    %
+    \begin{subfigure}[b]{.5\linewidth}
+        \centering
+        \includegraphics[width=1\linewidth]{images/eval/weights_BGMM_d128_k20.png}
+        \caption{D=128, K=20}
+        \label{fig:dblp_weights_d128_k20}
+    \end{subfigure}
+
+    \caption{Weights of communties generated by ComE BGMM+VI over hyperparameters representation sizes $D=[2, 128]$ and number of communities $K=[5, 10, 20]$.}
+    \label{fig:dblp_weights}
+\end{figure}
+
+Figure \ref{fig:dblp_weights} shows in subfigures \ref{fig:dblp_weights_d2_k5}, \ref{fig:dblp_weights_d2_k10}, and \ref{fig:dblp_weights_d2_k20}, that ComE BGMM+VI is able to identify $K=5$ communities for the initial hyperparameter settings $K=[5, 10, 20]$ if a two-dimensional embedding is choosen.
+For the 128-dimensional embeddings in subfigures \ref{fig:dblp_weights_d128_k5}, \ref{fig:dblp_weights_d128_k10}, and \ref{fig:dblp_weights_d128_k20}, ComE BGMM+VI is not able to drop communities to identify the $K=5$ with initial values of $K=[5,10,20]$.
+
+The finding from comparing community weights over the two- and 128-dimensional embedding of ComE BGMM+VI not dropping components for the 128-dimensional embeddings could explain the observation in Section \ref{sec:eval_nmi} of a better NMI for $D=2$ when $K=20$.
+
+It is left for future work to consider different weight concentration priors and further explorations into the ability of BGMM and variational inference to unused drop components dependant of the dimensionality of input space.
+
+\section{Results of Node Classification}
+
+Figure \ref{table:dblp_f1} shows the results of ComE GMM+EM and ComE BGMM+VI on the task node classification compared over multiple values for the hyperparameter number of communities $K$.
+The evaluation metrics micro- and macro-F1 used for node classification are described in Section \ref{sec:come_eval_metrics_nc}.
+
+First, embeddings are generated from the whole graph.
+Then, node classification is done by a support vector machine (SVM), which is trained on $90\%$ of the node embeddings and tested on the remaining $10\%$ of node embeddings.
+The two- or 128-dimensional node embeddings are used as features for the support vector machine.
+
+\begin{figure}[H]
+    \centering
+
+    \begin{subfigure}[b]{\linewidth}
+        \begin{table}[H]
+            \centering
+            \begin{tabular}{ l | l | c | c | c | c }
+                \hline
+                \empty                         & \#(communities) $K$ & 2             & 5             & 10            & 20            \\
+                \hline
+                \hline
+                \multirow{2}{*}{Micro-F1 (\%)} & ComE BGMM+VI        & \textbf{67.1} & 67.2          & 66.2          & \textbf{67.0} \\
+                                               & ComE                & 66.9          & \textbf{67.4} & \textbf{66.6} & 66.5          \\
+                \hline
+                \multirow{2}{*}{Macro-F1 (\%)} & ComE BGMM+VI        & \textbf{50.9} & 49.6          & 48.6          & \textbf{49.6} \\
+                                               & ComE                & 50.0          & \textbf{50.0} & \textbf{49.3} & 49.0          \\
+                \hline
+            \end{tabular}
+            \caption{two-dimensional embeddings.}
+            \label{table:dblp_f1_d2}
+        \end{table}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{\linewidth}
+        \begin{table}[H]
+            \centering
+            \begin{tabular}{ l | l | c | c | c | c }
+                \hline
+                \empty                         & \#(communities) $K$ & 2             & 5             & 10            & 20            \\
+                \hline
+                \hline
+                \multirow{2}{*}{Micro-F1 (\%)} & ComE BGMM+VI        & 92.6          & \textbf{92.6} & \textbf{93.3} & 92.3          \\
+                                               & ComE                & 92.6          & 92.5          & 93.2          & \textbf{92.7} \\
+                \hline
+                \multirow{2}{*}{Macro-F1 (\%)} & ComE BGMM+VI        & \textbf{92.7} & \textbf{92.4} & \textbf{93.0} & 91.9          \\
+                                               & ComE                & 92.6          & 92.3          & 92.9          & \textbf{92.7} \\
+                \hline
+            \end{tabular}
+            \caption{128-dimensional embeddings.}
+            \label{table:dblp_f1_d128}
+        \end{table}
+    \end{subfigure}
+
+    \caption{Micro- and Macro-F1 on the DBLP dataset with two- and 128-dimensional embeddings. Higher F1 values are better.}
+    \label{table:dblp_f1}
+\end{figure}
+
+The Micro- and Macro-F1 scores attained by ComE GMM+EM and ComE BGMM+VI node embeddings through an SVM are very similar.
+No advantage of one over the other can be determined from the results.
+The similarity between results for node classification could be due to the specialties of the used DBLP dataset.
+To gain more clarity, further datasets should be evaluated in future work.
+
+\subsection{Summary}
+
+Although a better conductance was achieved for large $K$ on the 128-dimensional embedding, a clear advantage of ComE BGMM+VI over ComE GMM+EM can not be identified for conductance over one experiment.
+
+For NMI, ComE BGMM+VI seems to perform slightly better than ComE when choosing a two-dimensional embedding instead of a 128-dimensional embedding and larger initial number of communities that the true number of communities: $K>K^*$.
+It is hypothesized that this advantage comes from the ability of BGMM+VI to drop unused components in lower-dimensional space.
+It appears that BGMM+VI does not drop components as easy in higher-dimensional space.
+
+Examining the weights of components produced by ComE BGMM+VI in two- and 128-dimensional space provides the insight, that ComE BGMM+VI produces the same distributions of 5 components' weights for $K=5$, $K=10$, and $K=20$ in two-dimensional space.
+In 128-dimensional space, on the other hand, ComE BGMM+VI does not drop components and remains with weight distributions including all the components added at initialization.
+In future work, it might be interesting to explore the weight concentration prior or other mechanisms of the BGMM+VI to allow ComE to drop unused communities at higher dimensionality as well.
+
+Evaluation of node classification aims to quantify the quality of classifications generated by node embeddings.
+Firstly, embeddings are generated for the whole graph.
+Then, a support vector machine (SVM) is trained with $90\%$ of the 128-dimensional node embeddings as features and true labels as target values.
+Ultimately, the micro- and macro-F1 scores are determined for the remaining $10\%$ testing data.
+In the experiment of the DBLP dataset, no advantage of using ComE BGMM+VI over GMM+EM was identified for node classification.
+It is reasonable to note, that node classification with micro- and macro-F1 of over $90\%$ is achieved, with the 128-dimensional embedding, leaving little room for optimization.
+
+A slight advantage of ComE BGMM+VI over ComE GMM+EM for community detection over the DBLP dataset under specific conditions was identified.
+No advantage for the task of node classification for ComE BGMM+VI over ComE GMM+EM could be identified for the DBLP dataset.
+
+It was identified, that while the BGMM with variational inference can drop unused components when operating in low dimensional space and achieve the same component weight distributions over multiple initial values for $K$, in high dimensional space, the BGMM does not drop components.
+
 %%% Bibliography

### 2020-10-14 paper-ComE-BGMM-VI.tex
write conclusion

@@
 In their original \citeyear{ComE} paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs}, \citeauthor{ComE} identify three evaluation tasks for experimentation: graph visualization, community detection and node classification \cite{ComE}.
-Community detection and node classification can objectivelly be quantitatively evaluated, while graph visualization is subjectively evaluated.
+Community detection and node classification can objectively be quantitatively evaluated, while graph visualization is subjectively evaluated.
 
@@
 
-\begin{figure}[H]
+\begin{figure*}[htbp]
     \centering
-    \begin{subfigure}{.5\linewidth}
+    \begin{subfigure}{.5\textwidth}
         \centering
@@
                     width=\linewidth,
-                    height=\linewidth,
+                    height=.5\linewidth,
                     legend style={at={(0.5,1)},
@@
     %
-    \begin{subfigure}{.5\linewidth}
+    \begin{subfigure}{.5\textwidth}
         \centering
@@
     \label{fig:dblp_conductance}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[htbp]
     \centering
@@
     \label{fig:dblp_nmi}
-\end{figure}
+\end{figure*}
 
@@
 Figure \ref{table:dblp_f1} shows the results of ComE GMM+EM and ComE BGMM+VI on the task node classification compared over multiple values for the hyperparameter number of communities $K$.
-The evaluation metrics micro- and macro-F1 used for node classification are described in Section \ref{sec:come_eval_metrics_nc}.
+The evaluation metrics micro- and macro-F1 are used.
 
@@
 
-For NMI, ComE BGMM+VI seems to perform slightly better than ComE when choosing a two-dimensional embedding instead of a 128-dimensional embedding and larger initial number of communities that the true number of communities: $K>K^*$.
+For NMI, ComE BGMM+VI seems to perform slightly better than ComE when choosing a two-dimensional embedding instead of a 128-dimensional embedding and a greater initial number of communities that the true number of communities: $K>5$.
 It is hypothesized that this advantage comes from the ability of BGMM+VI to drop unused components in lower-dimensional space.
@@
 In 128-dimensional space, on the other hand, ComE BGMM+VI does not drop components and remains with weight distributions including all the components added at initialization.
-In future work, it might be interesting to explore the weight concentration prior or other mechanisms of the BGMM+VI to allow ComE to drop unused communities at higher dimensionality as well.
 
-Evaluation of node classification aims to quantify the quality of classifications generated by node embeddings.
-Firstly, embeddings are generated for the whole graph.
-Then, a support vector machine (SVM) is trained with $90\%$ of the 128-dimensional node embeddings as features and true labels as target values.
-Ultimately, the micro- and macro-F1 scores are determined for the remaining $10\%$ testing data.
-In the experiment of the DBLP dataset, no advantage of using ComE BGMM+VI over GMM+EM was identified for node classification.
+In the experiment on the DBLP dataset, no advantage of using ComE BGMM+VI over GMM+EM was identified for node classification by SVM.
 It is reasonable to note, that node classification with micro- and macro-F1 of over $90\%$ is achieved, with the 128-dimensional embedding, leaving little room for optimization.
@@
 
+\section{Conclusion}
+
+ComE BGMM+VI, an extension of the community embedding algorithm ComE is presented in this paper.
+ComE BGMM+VI utilizes a Bayesian Gaussian mixture model instead of a non-Bayesian Gaussian mixture model for community embedding and detection.
+The Bayesian community embedding algorithm ComE BGMM+VI is compared to ComE GMM+EM visually on the Karate Club graph dataset and quantitatively evaluated on community detection and node classification the DBLP dataset.
+
+Visually comparing ComE BGMM+VI to ComE GMM+EM clearly shows the advantage of ComE BGMM+VI over ComE GMM+EM for an unknown number of communities $K$ and a low representation size $D$.
+ComE BGMM+VI is able to drop unused communities and reduce $K$ to the number of communities present at the resolution captured at the chosen representation size $D$ through the weight concentration prior hyperparameter trade-off parameter.
+
+A slight edge of ComE BGMM+VI over the original ComE as identified for community detection NMI in a low-dimensional embedding for $K$ larger than the true number of communities.
+More notably, ComE BGMM+VI captures 5 communities in the DBLP dataset at representation size $D=2$ for all tested values for the number of communities $K>5$.
+The true number of communities in the DBLP dataset is 5.
+ComE with non-Bayesian Gaussian mixture model is not able to drop communities for $K>5$, as seen in Figure \ref{fig:dblp_weights}.
+
 %%% Bibliography

### 2022-05-19 paper-ComE-BGMM-VI.tex
move visual comparison K=15 to first page
sort some figures
👨‍💻 add todos

@@
 \emergencystretch=8em
+
+\newcommand{\visresults}[2]{
+    \begin{figure}[H]
+        \centering
+        \begin{subfigure}{.15\textwidth}
+            \centering
+            \includegraphics[width=1\linewidth]{images/results/#1/graph_#2.png}
+            \caption{Graph}
+        \end{subfigure}%
+        %
+        \begin{subfigure}{.15\textwidth}
+            \centering
+            \includegraphics[width=1\linewidth]{images/results/#1/embeddings_#2.png}
+            \caption{Embeddings}
+        \end{subfigure}%
+        %
+        \begin{subfigure}{.15\textwidth}
+            \centering
+            \includegraphics[width=1\linewidth]{images/results/#1/weights_#2.png}
+            \caption{Weights}
+        \end{subfigure}%
+        %
+        \caption{ComE with #1, $K=#2$}
+        \label{fig:results#1#2}
+    \end{figure}
+}
 
@@
 
+\visresults{GMM}{15}
+\visresults{BGMM}{15}
+
 The recent \citeyear{ComE} graph embeddings algorithm ComE is extended similarily to the \citeyear{ComE+} ComE+ by taking a Bayesian approach.
@@
 On part of comparing the Bayesian Gaussian mixture model with variational inference to the Gaussian mixture model with expectation-maximization, the \citeyear{Bishop06} book \textit{Pattern Recognition and Machine Learning (Information Science and Statistics)} by \citeauthor{Bishop06} includes essential statistics and information science knowledge and explanations \cite{Bishop06}.
+
+\section{Work Proceeding}
+
+TODO programming language, frameworks, and graphing libraries
+It is also worth pointing out the libraries of the data processing language with which calculations were made and graphs were built.
+Programming/application software development (libraries for graphs).
 
@@
 
-\begin{table}[htbp]
+\begin{table}[H]
     \centering
@@
 
+TODO hyperparameters GMM
+In the article, it is desirable to state in more detail the essence of the work of algorithms, to describe hyperparameters in more detail (where do they come from?)
+Where are the hyper parameters from? Node2Vec, EM, VI, ...
+
 \subsection{Non-Bayesian and Bayesian Gaussian mixture models}
 
-\begin{figure}[htbp]
+\begin{figure}[H]
     \centering
@@
 \end{table}
+
+TODO hyperparameters BGMM
+In the article, it is desirable to state in more detail the essence of the work of algorithms, to describe hyperparameters in more detail (where do they come from?)
+Where are the hyper parameters from? Node2Vec, EM, VI, ...
 
@@
 
-\newcommand{\visresults}[2]{
-    \begin{figure}[H]
-        \centering
-        \begin{subfigure}{.15\textwidth}
-            \centering
-            \includegraphics[width=1\linewidth]{images/results/#1/graph_#2.png}
-            \caption{Graph}
-        \end{subfigure}%
-        %
-        \begin{subfigure}{.15\textwidth}
-            \centering
-            \includegraphics[width=1\linewidth]{images/results/#1/embeddings_#2.png}
-            \caption{Embeddings}
-        \end{subfigure}%
-        %
-        \begin{subfigure}{.15\textwidth}
-            \centering
-            \includegraphics[width=1\linewidth]{images/results/#1/weights_#2.png}
-            \caption{Weights}
-        \end{subfigure}%
-        %
-        \caption{ComE with #1, $K=#2$}
-        \label{fig:results#1#2}
-    \end{figure}
-}
-
 \subsection{K=2}
@@
 
-\visresults{GMM}{15}
-\visresults{BGMM}{15}
+%\visresults{GMM}{15}
+%\visresults{BGMM}{15}
 

### 2022-05-21 paper-ComE-BGMM-VI.tex
Implement Feedback from 18.05.2022

--- 2020-10-14 paper-ComE-BGMM-VI.tex
@@
 
+\visresults{GMM}{15}
+\visresults{BGMM}{15}
+
 The recent \citeyear{ComE} graph embeddings algorithm ComE is extended similarily to the \citeyear{ComE+} ComE+ by taking a Bayesian approach.
 The open-source code for the Bayesian approach to ComE's community embedding is published on GitHub\footnote{at \url{https://github.com/abegehr/ComE_BGMM}} and serves as a contribution to community embedding research \cite{ComE_BGMM_GH}.
+The source code is written in Python and tested with Python 3.6.10.
+For node and context embedding, ComE's training\_sdg\_inner Cython binary is utilized \cite{van1995python, ComE_GH}.
 
@@
 
-Multiple surveys and articles on graph embeddings were consulted to build a full picture of the current state of graph embedding research.
+Surveys and articles on graph embeddings were consulted to build a full picture of the current state of graph embedding research.
 Especially the \citeyear{Goyal_2018} survey \textit{Graph Embedding Techniques, Applications, and Performance: A Survey} by \citeauthor{Goyal_2018} and the \citeyear{rossi20tkdd-roles} paper \textit{On Proximity and Structural Role-based Embeddings in Networks: Misconceptions, Techniques, and Applications} by \citeauthor{rossi20tkdd-roles} have proven to be primary resources for understanding the current landscape of graph embedding research \cite{Goyal_2018, rossi20tkdd-roles}.
@@
 On part of comparing the Bayesian Gaussian mixture model with variational inference to the Gaussian mixture model with expectation-maximization, the \citeyear{Bishop06} book \textit{Pattern Recognition and Machine Learning (Information Science and Statistics)} by \citeauthor{Bishop06} includes essential statistics and information science knowledge and explanations \cite{Bishop06}.
+
+During implementation, data transformation, testing, and plotting the following libraries and frameworks have proved invaluable: NumPy, Pandas, Matplotlib, NetworkX, and Jupyter Notebooks \cite{harris2020array, mckinney-proc-scipy-2010, reback2020pandas, Hunter:2007, SciPyProceedings_11, Kluyver2016jupyter}.
 
@@
         \hline
-        alpha                & $\alpha$ & Trade-off parameter for context embedding                                       \\
+        alpha                & $\alpha$ & trade-off parameter for context embedding                                       \\
         \hline
-        beta                 & $\beta$  & Trade-off parameter for community embedding                                     \\
+        beta                 & $\beta$  & trade-off parameter for community embedding                                     \\
         \hline
@@
 
+The hyperparameters number\_walks, walk\_length, and representation\_size are utilised in the node embedding step Node2Vec to achieve an initial node embedding optimizing for first and second oder proximity.
+reg\_covar,  batch\_size, window\_size, negative, and lr are utilized in the context embedding step.
+num\_workers is a performance parameter which allows multithreading the node embedding or context embedding steps.
+num\_iter, and com\_n\_init control the number of iterations to run over ComE's closed loop algorithm and how many random initialisations of the community embedding to try out, respectivelly.
+alpha and beta allow adjusting the weighting between context and community embedding.
+Communities $K$ is the number of Gaussian generators as pertained to the Gaussian Mixture Model.
+
 \subsection{Non-Bayesian and Bayesian Gaussian mixture models}
 
-\begin{figure}[htbp]
+\begin{figure}[H]
     \centering

### 2022-05-21 paper-ComE-BGMM-VI.tex
🧹 cleanup

@@
 
-The advantage of a smaller embedding becomes clearer for larger values of $K>K^*$.
+The potential advantage of a smaller embedding becomes clearer for larger values of $K>K^*$.
 For $K=20$ the NMI for the 128-dimensional embedding is around $0.3$, while the BGMM two-dimensional embedding manages to hold at $0.45$.
@@
 
-To understand why ComE BGMM+VI is able to maintain an NMI score for larger values of $K$ for the two-dimensional embedding, but not for the 128-dimensional embedding, the weights of the components for $K=20$ is presented:
-
-\begin{figure}[H]
+\begin{figure}
     \centering
